---
title: "EpiParam"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    navbar:
    - icon: "fa fa-rss"
      href: http://onehealth.or.kr/
      align: right
    vertical_layout: fill
    theme:
      version: 4
      fg: '#030303'
      bg: '#FBFBFB'
      navbar-bg: '#8B5A00'
      sidebar-bg: '#8B5A00'
      smart: no
runtime: shiny
---

```{=html}
<style type="text/css">

:root {
--epiparam-font: Jost, "Segoe UI", Roboto, "Helvetica Neue", Arial, sans-serif;
}

.vis-network .vis-export {
position: absolute !important;
top: 10px !important;
right: 10px !important;
z-index: 1000 !important;
background-color: #4169E1 !important;
color: white !important;
border: none !important;
border-radius: 4px !important;
padding: 6px 12px !important;
font-weight: bold !important;
margin-right: 0 !important;
}

.vis-network .vis-export:hover {
background-color: #3151A3 !important;
}

.irs--shiny .irs-min, .irs--shiny .irs-max {
top: 0;
padding: 1px 3px;
text-shadow: none;
background-color: rgba(25, 25, 26, 0.1);
border-radius: 3px;
font-size: 14px;
line-height: 1.333;
}

.irs--shiny .irs-from, .irs--shiny .irs-to, .irs--shiny .irs-single {
color: #FBFBFB;
text-shadow: none;
padding: 1px 3px;
background-color: #000000;
border-radius: 3px;
font-size: 14px;
line-height: 1.333;
}

.irs--shiny .irs-bar {
top: 25px;
height: 8px;
border-top: 1px solid #70b2e0;
border-bottom: 1px solid #70b2e0;
background: #70b2e0;
cursor: s-resize;
z-index: 2;
}

.error-highlight {
border: 2px solid #ff0000 !important;
background-color: #ffeeee !important;
}

body hr {
border-color: #ffffff;
margin: 10px 0;
border: 2px solid #ffffff !important;
}

.rt-bordered .rt-table, .rt-outlined .rt-table {
border-style: solid;
border-top-left-radius: 1.25rem;
}

h4, .h4 {
font-size: 10px;
}

.chart-wrapper .chart-stage {
color: #000000;
padding: 0px;
}

.navbar.navbar-inverse {
background-color: #8B5A00 !important;
border-bottom-right-radius: 5.25rem;
}

.navbar-brand, .navbar.navbar-inverse .navbar-brand {
color: white !important;
}

.navbar-brand:hover, .navbar.navbar-inverse .navbar-brand:hover {
color: white !important;
}

.nav-tabs-custom > .nav-tabs > li.active {
border-top-color: #8B5A00;
}

.navbar.navbar-inverse ul.nav.navbar-nav>li {
color: #fff;
font-size: 16px;
font-family: var(--epiparam-font);
}

.nav-tabs-custom > .nav-tabs {
margin: 0;
border-bottom: 1px solid #8B5A00;
color: white;
border-top-right-radius: 10px;
border-top-left-radius: 3px;
font-family: var(--epiparam-font);
}

.nav-tabs-custom > .nav-tabs > li > a, .nav-tabs-custom > .nav-tabs > li > a:active {
color: #000000;
font-size: 0.9375rem;
border-radius: 0;
padding: 3px 10px 5px;
text-transform: none;
font-weight: 500;
}


.chart-wrapper, .nav-tabs-custom, .sbframe-commentary {
background: #FFFFFF;
border: 1px solid #B2B2B2;
border-radius: 0.25rem;
margin-bottom: 8px;
margin-right: 8px;
}

.navbar-brand {
font-family: var(--epiparam-font);
font-weight: 700;
font-size: 38px;
padding: 0px;
}

.chart-title {
border-bottom: 1px solid #B7B7B7;
font-size: 0.9375rem;
text-align: center;
font-weight: bold;
padding: 7px 10px 4px;
}

.section.sidebar {
font-family: var(--epiparam-font);
font-size: 14px;
background-color: #8B5A00 !important;
color: white;
padding: 0px;
border: 1px solid #8B5A00;
position: fixed;
top: 62px;
left: 0;
bottom: 0;
border-right: 1px solid #dee2e6;
background-color: rgba(39, 128, 227, 0.1);
padding-top: 10px;
padding-left: 10px;
padding-right: 10px;
visibility: hidden;
overflow: auto;
}

/* dataset preview tabs hold the data check message and the table under it.
flexdashboard stretches every output to the full height of the tab, which
pushed the table out of sight; here each keeps its own height and the tab
scrolls instead */
.data-preview .chart-shim {
overflow-y: auto;
}

.data-preview .chart-stage .shiny-bound-output {
height: auto !important;
}

/* one line per row; wide files scroll sideways */
.data-preview table.dataTable th,
.data-preview table.dataTable td {
white-space: nowrap;
}

/* the logo sits below the last control of the sidebar, at the bottom of the
window when there is room for it, so it never covers a control */
.section.sidebar > form {
display: flow-root;
position: relative;
min-height: 100%;
padding-bottom: 80px;
}

.section.sidebar .shiny-image-output {
position: absolute;
left: 0;
bottom: 10px;
width: auto !important;
height: auto !important;
}

a {
color: #28282b;
font-size: 19px;
font-family: var(--epiparam-font);
background-color: transparent;
}

.nav-tabs .nav-link.active, .nav-tabs>li>a.active, .nav-tabs .nav-pills>li>a.active, .nav-tabs :where(ul.nav.navbar-nav > li)>a.active, .nav-tabs .nav-item.show .nav-link, .nav-tabs .nav-item.in .nav-link, .nav-tabs .nav-item.show .nav-tabs>li>a, .nav-tabs .nav-item.in .nav-tabs>li>a, .nav-tabs .nav-item.show .nav-pills>li>a, .nav-tabs .nav-item.in .nav-pills>li>a, .nav-tabs>li.show .nav-link, .nav-tabs>li.in .nav-link, .nav-tabs>li.show .nav-tabs>li>a, .nav-tabs>li.in .nav-tabs>li>a, .nav-tabs>li.show .nav-pills>li>a, .nav-tabs>li.in .nav-pills>li>a, .nav-tabs .nav-pills>li.show .nav-link, .nav-tabs .nav-pills>li.in .nav-link, .nav-tabs .nav-pills>li.show .nav-tabs>li>a, .nav-tabs .nav-pills>li.in .nav-tabs>li>a, .nav-tabs .nav-pills>li.show .nav-pills>li>a, .nav-tabs .nav-pills>li.in .nav-pills>li>a, .nav-tabs .nav-item.show :where(ul.nav.navbar-nav > li)>a, .nav-tabs .nav-item.in :where(ul.nav.navbar-nav > li)>a, .nav-tabs>li.show :where(ul.nav.navbar-nav > li)>a, .nav-tabs>li.in :where(ul.nav.navbar-nav > li)>a, .nav-tabs .nav-pills>li.show :where(ul.nav.navbar-nav > li)>a, .nav-tabs .nav-pills>li.in :where(ul.nav.navbar-nav > li)>a, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-link, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-link, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-tabs>li>a, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-tabs>li>a, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-pills>li>a, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-pills>li>a, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) :where(ul.nav.navbar-nav > li)>a, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) :where(ul.nav.navbar-nav > li)>a {
color: #b814aa;
background-color: #FBFBFB;
border-color: #B7B7B7 #B7B7B7 #FBFBFB;
}

.chart-wrapper {
width: 100% !important;
}

pre {
font-family: var(--epiparam-font);
font-size: medium;
}

.btn-primary {
color: #FBFBFB;
background-color: #ff7100;
border-color: #ff7100;
}


</style>

```

```{r setup-chunk, include=FALSE}
# load libraries
library(shinyjs)
library(htmltools)
library(knitr)
library(flexdashboard)
library(lubridate)
library(bslib)
library(shiny)
library(tidyverse)
library(janitor)
library(plotly)
library(MASS)
library(R0)
library(DT)
library(EpiEstim)
library(incidence2)
library(visNetwork)
library(webshot)
library(webshot2)
library(officer)
library(networkD3)
library(highcharter)
library(fitdistrplus)
library(stringr)
library(rstan)
library(loo)

# suppress browser redirect on Stan warnings
options(warn = -1) 
rstan_options(auto_write = TRUE)

# the mcmc sampler starts from this seed, so the same data and settings give the
# same estimates on every run. change it to check that a result does not depend
# on the seed.
STAN_SEED <- 20260423

# date parsing function
parse_date <- function(x) {
  x <- trimws(as.character(x))
  parsed <- lubridate::parse_date_time(x, orders = c("dmy", "ymd", "mdy"), quiet = TRUE)
  as.Date(parsed)
}

# --- input checks ----------------------------------------------------------
# rows the app cannot use were previously dropped without a word, so a typo in a
# date or a transmission link that points at no one looked exactly like clean
# data. every page now marks each row with what became of it, shows the panel
# below above its preview, and previews the parsed values rather than the raw
# text of the file.

# values that were supplied but could not be read as a date
unreadable_values <- function(raw, parsed) {
  raw <- trimws(as.character(raw))
  blank <- !nzchar(raw) | is.na(raw) | toupper(raw) %in% c("NA", "N/A", "NULL", "-")
  unique(raw[!blank & is.na(parsed)])
}

# turns a status column into one readable line per problem, with examples
problems_from_status <- function(status, examples = list()) {
  bad <- status[status != "used"]
  if (!length(bad)) return(list())
  tab <- sort(table(bad), decreasing = TRUE)
  lapply(names(tab), function(k) {
    n <- as.integer(tab[[k]])
    txt <- sprintf("%d row%s: %s", n, if (n == 1) "" else "s", sub("^excluded: ", "", k))
    ex <- examples[[k]]
    if (!is.null(ex) && length(ex)) {
      txt <- paste0(txt, " (for example: ",
                    paste0("\u201c", utils::head(ex, 3), "\u201d", collapse = ", "), ")")
    }
    txt
  })
}

# the panel shown above each dataset preview
data_check_panel <- function(n_read, n_used, problems = list(),
                             used_label = "used in the analysis",
                             out_label  = "left out") {
  n_out <- n_read - n_used
  good  <- n_out == 0 && !length(problems)
  htmltools::div(
    style = paste0(
      "border-left:4px solid ", if (good) "#2E7D32" else "#B26A00", ";",
      "background:", if (good) "#EDF7ED" else "#FFF4E5", ";",
      "padding:10px 14px;margin:6px 0 12px 0;border-radius:4px;",
      "font-size:14px;line-height:1.5;color:#222222;"
    ),
    htmltools::tags$b(sprintf(
      "%d row%s in the current selection, %d %s, %d %s.",
      n_read, if (n_read == 1) "" else "s", n_used, used_label, n_out, out_label)),
    if (length(problems)) {
      htmltools::tags$ul(style = "margin:6px 0 0 0;padding-left:18px;",
                         lapply(problems, htmltools::tags$li))
    } else {
      htmltools::div(style = "margin-top:2px;", "Every row could be read.")
    },
    htmltools::div(
      style = "margin-top:6px;font-size:13px;color:#555555;",
      "The table below shows the values as the app read them. The status column says what happened to each row."
    )
  )
}

# helper functions for model fitting (MLE pages)
fit_parametric_model <- function(data, dist, shift_val = 0) {
  data_fit <- data + shift_val
  
  fit <- tryCatch({
    MASS::fitdistr(data_fit, dist)
  }, error = function(e) return(NULL))
  
  if(is.null(fit)) return(NULL)
  
  est <- fit$estimate
  
  if (dist == "lognormal") {
    mean_val <- exp(est["meanlog"] + est["sdlog"]^2 / 2)
    var_val <- (exp(est["sdlog"]^2) - 1) * exp(2 * est["meanlog"] + est["sdlog"]^2)
    sd_val <- sqrt(var_val)
    ci_95 <- qlnorm(0.95, est["meanlog"], est["sdlog"])
    
    raw_params <- data.frame(
      parameter = c("estimate.meanlog", "estimate.sdlog", "sd.meanlog", "sd.sdlog"),
      value = round(c(est["meanlog"], est["sdlog"], fit$sd["meanlog"], fit$sd["sdlog"]), 2)
    )
    
  } else if (dist == "weibull") {
    mean_val <- est["scale"] * gamma(1 + 1/est["shape"])
    var_val <- est["scale"]^2 * (gamma(1 + 2/est["shape"]) - (gamma(1 + 1/est["shape"]))^2)
    sd_val <- sqrt(var_val)
    ci_95 <- qweibull(0.95, est["shape"], est["scale"])
    
    raw_params <- data.frame(
      parameter = c("estimate.shape", "estimate.scale", "sd.shape", "sd.scale"),
      value = round(c(est["shape"], est["scale"], fit$sd["shape"], fit$sd["scale"]), 2)
    )
    
  } else if (dist == "gamma") {
    mean_val <- est["shape"] / est["rate"]
    var_val <- est["shape"] / (est["rate"]^2)
    sd_val <- sqrt(var_val)
    ci_95 <- qgamma(0.95, est["shape"], est["rate"])
    
    raw_params <- data.frame(
      parameter = c("estimate.shape", "estimate.rate", "sd.shape", "sd.rate"),
      value = round(c(est["shape"], est["rate"], fit$sd["shape"], fit$sd["rate"]), 2)
    )
    
  } else if (dist == "normal") {
    mean_val <- est["mean"]
    sd_val <- est["sd"]
    ci_95 <- qnorm(0.95, mean_val, sd_val)
    var_val <- sd_val^2
    
    raw_params <- data.frame(
      parameter = c("estimate.mean", "estimate.sd", "sd.mean", "sd.sd"),
      value = round(c(est["mean"], est["sd"], fit$sd["mean"], fit$sd["sd"]), 2)
    )
  }
  
  # a skewed distribution fitted to data moved to the right is a shifted
  # distribution: report the shift next to its parameters, and put the moments
  # back on to the observed scale
  if (shift_val != 0) {
    raw_params <- rbind(raw_params,
                        data.frame(parameter = "shift.applied",
                                   value = round(shift_val, 2)))
  }

  final_mean <- as.numeric(mean_val) - shift_val
  final_var <- as.numeric(var_val)
  final_ci <- as.numeric(ci_95) - shift_val
  
  list(
    model = toupper(dist),
    fit_obj = fit,
    shift = shift_val,
    raw_params_df = raw_params,
    moments = c(mean=final_mean, var=final_var, ci95=final_ci),
    aic = AIC(fit),
    bic = BIC(fit)
  )
}

# helper to generate plot data for incubation (MLE page)
get_plot_data_inc <- function(data, fit_res) {
  h_breaks <- seq(floor(min(data))-0.5, ceiling(max(data))+0.5, by = 1)
  hist_data <- hist(data, plot = FALSE, breaks = h_breaks)
  
  hist_df <- data.frame(
    from = hist_data$breaks[-length(hist_data$breaks)],
    to = hist_data$breaks[-1],
    y = hist_data$density,
    x = hist_data$mids
  )
  
  est <- fit_res$fit_obj$estimate
  d_name <- tolower(fit_res$model)
  min_x_calc <- max(0.01, min(data) - 1.5)
  x_vals <- seq(min_x_calc, max(data) + 1.5, length.out = 500)
  
  suppressWarnings({
    y_vals <- tryCatch({
      switch(d_name,
             "lognormal" = dlnorm(x_vals, meanlog = est["meanlog"], sdlog = est["sdlog"]),
             "weibull" = dweibull(x_vals, shape = est["shape"], scale = est["scale"]),
             "gamma"  = dgamma(x_vals, shape = est["shape"], rate = est["rate"])
      )
    }, error = function(e) return(rep(0, length(x_vals))))
  })
  
  curve_data <- data.frame(x = x_vals, y = y_vals)
  list(hist = hist_df, curve = curve_data)
}

# helper to generate plot data for serial interval (MLE page)
get_plot_data_si <- function(data, fit_res) {
  h_breaks <- seq(floor(min(data))-1, ceiling(max(data))+1, by = 2) 
  hist_data <- hist(data, plot = FALSE, breaks = h_breaks)
  hist_df <- data.frame(from = hist_data$breaks[-length(hist_data$breaks)], to = hist_data$breaks[-1], y = hist_data$density, x = hist_data$mids)
  
  # the shift the model was actually fitted with, so the curve lines up with
  # the histogram of the observed values
  shift_val <- if (!is.null(fit_res$shift)) fit_res$shift else
                 if (tolower(fit_res$model) == "normal") 0 else 1
  est <- fit_res$fit_obj$estimate
  d_name <- tolower(fit_res$model)
  x_vals <- seq(min(data) - 5, max(data) + 5, length.out = 500)
  x_trans <- x_vals + shift_val
  
  if(shift_val > 0) x_trans[x_trans <= 0] <- NA
  
  suppressWarnings({
    y_vals <- tryCatch({
      switch(d_name,
             "lognormal" = dlnorm(x_trans, meanlog = est["meanlog"], sdlog = est["sdlog"]),
             "weibull"   = dweibull(x_trans, shape = est["shape"], scale = est["scale"]),
             "gamma"     = dgamma(x_trans, shape = est["shape"], rate = est["rate"]),
             "normal"    = dnorm(x_trans, mean = est["mean"], sd = est["sd"])
      )
    }, error = function(e) return(rep(0, length(x_vals))))
  })
  y_vals[is.na(y_vals)] <- 0
  curve_data <- data.frame(x = x_vals, y = y_vals)
  list(hist = hist_df, curve = curve_data)
}

# ---------------------------------------------------------------------------
# bayesian models: stan source and compilation
#
# the stan code below is the only definition of the four bayesian models. a
# compiled stan model is a shared library that loads only on the platform, R
# version and rstan version that built it, so a binary built elsewhere fails
# with "the compiled object from C++ code for this model is invalid".
#
# a model is taken from, in this order:
#   1. this R session, once it has been loaded
#   2. the per-user cache (see stan_cache_dir), filled by an earlier compilation
#      on this computer
#   3. the stan_models folder of the app, filled by precompile_models.R --bundle
#      on a computer with the same platform, R version and rstan version
#   4. a compilation on this computer, in a separate R process. that process
#      starts when the app opens and works through the models one at a time,
#      the one the user is waiting for first, so the first run seldom waits for
#      the whole compilation and the app shows how long it has been going.
#
# a binary is used only when it was built for this platform and from the source
# in get_stan_code(), so the code shown here and the code that produced the
# results can never drift apart. compiling takes one to three minutes and about
# 3.5 GB of memory per model, which is why only one model is compiled at a time.
# ---------------------------------------------------------------------------

# --- stan helpers: start ------------------------------------------------
# precompile_models.R and the background compilation both read the block
# between these two markers, so the models stay defined in exactly one place.
# please keep them in place.

STAN_MODELS <- c("Lognormal", "Weibull", "Gamma", "Normal")

# bayesian helper: the stan source each model is built from
get_stan_code <- function(model_name = "Weibull") {

  if (model_name == "Weibull") {
    return("
data {
  int<lower=0> N;
  vector[N] tStartExposure;
  vector[N] tEndExposure;
  vector[N] tSymptomOnset;
}
parameters {
  real<lower=0> alphaInc; // shape
  real<lower=0> sigmaInc; // scale
  vector<lower=0, upper=1>[N] uE;
}
transformed parameters {
  vector[N] tE;
  tE = tStartExposure + uE .* (tEndExposure - tStartExposure);
}
model {
  // priors
  alphaInc ~ lognormal(log(2.0), 1.0);
  sigmaInc ~ lognormal(log(3.0), 1.0);

  // likelihood
  target += weibull_lpdf((tSymptomOnset - tE) + 0.1 | alphaInc, sigmaInc);
}
")
  }
  else if (model_name == "Gamma") {
    return("
data {
  int<lower=0> N;
  vector[N] tStartExposure;
  vector[N] tEndExposure;
  vector[N] tSymptomOnset;
}
parameters {
  real<lower=0> alphaInc; // shape
  real<lower=0> sigmaInc; // rate
  vector<lower=0, upper=1>[N] uE;
}
transformed parameters {
  vector[N] tE;
  tE = tStartExposure + uE .* (tEndExposure - tStartExposure);
}
model {
  // priors
  alphaInc ~ lognormal(log(2.0), 1.0);
  sigmaInc ~ lognormal(log(0.66), 1.0);

  // likelihood
  target += gamma_lpdf((tSymptomOnset - tE) + 0.1 | alphaInc, sigmaInc);
}
")
  }
  else if (model_name == "Lognormal") {
    return("
data {
  int<lower=0> N;
  vector[N] tStartExposure;
  vector[N] tEndExposure;
  vector[N] tSymptomOnset;
}
parameters {
  real alphaInc;          // meanlog
  real<lower=0> sigmaInc; // sdlog
  vector<lower=0, upper=1>[N] uE;
}
transformed parameters {
  vector[N] tE;
  tE = tStartExposure + uE .* (tEndExposure - tStartExposure);
}
model {
  // priors
  alphaInc ~ normal(log(3.0), 1.0);
  sigmaInc ~ normal(0, 1.0);

  // likelihood
  target += lognormal_lpdf((tSymptomOnset - tE) + 0.1 | alphaInc, sigmaInc);
}
")
  }
  else if (model_name == "Normal") {
    return("
data {
  int<lower=0> N;
  vector[N] tStartExposure;
  vector[N] tEndExposure;
  vector[N] tSymptomOnset;
}
parameters {
  real alphaInc;          // mean
  real<lower=0> sigmaInc; // sd
  vector<lower=0, upper=1>[N] uE;
}
transformed parameters {
  vector[N] tE;
  tE = tStartExposure + uE .* (tEndExposure - tStartExposure);
}
model {
  // priors
  alphaInc ~ normal(3.0, 1.0);
  sigmaInc ~ normal(0, 1.0);

  // likelihood
  target += normal_lpdf((tSymptomOnset - tE) + 0.1 | alphaInc, sigmaInc);
}
")
  }
  return(NULL)
}

# where the models compiled on this computer are kept. set EPIPARAM_STAN_CACHE
# to place them somewhere else, for example on a shared server installation.
stan_cache_dir <- function() {
  d <- Sys.getenv("EPIPARAM_STAN_CACHE", "")
  if (!nzchar(d)) {
    d <- tryCatch(tools::R_user_dir("EpiParam", "cache"),
                  error = function(e) file.path(tempdir(), "EpiParam-stan"))
  }
  dir.create(d, recursive = TRUE, showWarnings = FALSE)
  d
}

# the file name carries the platform, the rstan version and the R version, so a
# binary is never handed to an installation that cannot load it. R keeps
# compiled code compatible within a minor version (4.4.0 to 4.4.3, say), so the
# R version is given as major.minor.
stan_model_file_name <- function(model_name) {
  r_minor <- strsplit(R.version$minor, ".", fixed = TRUE)[[1]][1]
  sprintf("%s-%s-rstan%s-R%s.%s.rds", tolower(model_name), R.version$platform,
          utils::packageVersion("rstan"), R.version$major, r_minor)
}

stan_cache_file <- function(model_name) {
  file.path(stan_cache_dir(), stan_model_file_name(model_name))
}

# models built in advance with precompile_models.R --bundle and shipped in the
# stan_models folder next to the app
stan_bundle_file <- function(model_name) {
  file.path("stan_models", stan_model_file_name(model_name))
}

# stan source reduced to what matters for a comparison: rstan normalises the
# code it stores (it drops blank lines), so the two are compared without blank
# lines or trailing spaces rather than character by character
stan_code_key <- function(x) {
  x <- gsub("\r\n", "\n", as.character(x), fixed = TRUE)
  lines <- trimws(strsplit(x, "\n", fixed = TRUE)[[1]])
  paste(lines[nzchar(lines)], collapse = "\n")
}

# a compiled object is safe to use only if it was built for this platform and
# from the source the app is asking for
stan_model_usable <- function(mod, code) {
  if (!inherits(mod, "stanmodel")) return(FALSE)
  if (!identical(stan_code_key(mod@model_code), stan_code_key(code))) return(FALSE)
  built_on <- tryCatch(mod@dso@system, error = function(e) NA_character_)
  if (!identical(built_on, paste(R.version$arch, R.version$os, sep = ", "))) return(FALSE)
  # rstan's own check as well, where it is available: it also catches a dso
  # that is no longer loaded
  isTRUE(tryCatch(rstan:::is_sm_valid(mod), error = function(e) TRUE))
}

# the model saved in a file, or NULL when there is none this installation can
# use. a cached file that cannot be used (older source, another installation)
# is removed so that it gets rebuilt.
stan_read_model <- function(path, code, remove_stale = FALSE) {
  if (!file.exists(path)) return(NULL)
  mod <- tryCatch(readRDS(path), error = function(e) NULL)
  if (stan_model_usable(mod, code)) return(mod)
  if (remove_stale) unlink(path)
  NULL
}

# a model being compiled is marked by a folder next to its cache file. creating
# a folder either succeeds or fails as a whole, so two R processes never both
# start compiling the same model.
stan_lock_dir <- function(model_name) paste0(stan_cache_file(model_name), ".lock")

# the compiler's message from a failed attempt, kept for the app to show
stan_error_file <- function(model_name) paste0(stan_cache_file(model_name), ".error")

# is an R process on this computer compiling the model right now? a lock left
# behind by a process that stopped part way is removed.
stan_lock_busy <- function(model_name) {
  lock <- stan_lock_dir(model_name)
  if (!dir.exists(lock)) return(FALSE)
  age <- as.numeric(difftime(Sys.time(), file.mtime(lock), units = "secs"))
  pid <- tryCatch(as.integer(readLines(file.path(lock, "pid"), warn = FALSE)[1]),
                  error = function(e) NA_integer_, warning = function(w) NA_integer_)
  busy <- if (is.na(pid)) {
    age < 30                                   # just created, pid not written yet
  } else if (requireNamespace("ps", quietly = TRUE)) {
    age < 1800 && isTRUE(tryCatch(ps::ps_is_running(ps::ps_handle(pid)),
                                  error = function(e) FALSE))
  } else {
    age < 900
  }
  if (!busy) unlink(lock, recursive = TRUE)
  busy
}

stan_lock_take <- function(model_name) {
  if (stan_lock_busy(model_name)) return(FALSE)
  lock <- stan_lock_dir(model_name)
  if (!dir.create(lock, showWarnings = FALSE)) {
    # either another process got there first, or the cache folder cannot be
    # written to, in which case compile anyway
    return(!dir.exists(lock))
  }
  writeLines(as.character(Sys.getpid()), file.path(lock, "pid"))
  TRUE
}

# compile a model into the cache, unless a usable copy exists already or
# another process is compiling it. run by the background process and by
# precompile_models.R. returns TRUE when a usable model is in place.
stan_build <- function(model_name) {
  code  <- get_stan_code(model_name)
  cache <- stan_cache_file(model_name)
  if (!is.null(stan_read_model(cache, code, remove_stale = TRUE)) ||
      !is.null(stan_read_model(stan_bundle_file(model_name), code))) {
    return(invisible(TRUE))
  }
  if (!stan_lock_take(model_name)) return(invisible(FALSE))
  on.exit(unlink(stan_lock_dir(model_name), recursive = TRUE), add = TRUE)
  unlink(stan_error_file(model_name))
  message("compiling the ", tolower(model_name), " model for this computer")
  mod <- tryCatch(
    rstan::stan_model(model_code = code,
                      model_name = paste0("epiparam_", tolower(model_name))),
    error = function(e) {
      # rstan's own message says little; the end of the compiler output, which
      # the background process writes to a log, says what went wrong
      log <- Sys.getenv("EPIPARAM_STAN_LOG")
      out <- if (nzchar(log) && file.exists(log)) tail(readLines(log, warn = FALSE), 15)
      try(writeLines(c(conditionMessage(e), out), stan_error_file(model_name)), silent = TRUE)
      NULL
    }
  )
  if (is.null(mod)) return(invisible(FALSE))
  # saved under a temporary name first, so no one ever reads a half-written file
  part <- paste0(cache, ".part", Sys.getpid())
  saved <- !inherits(try(saveRDS(mod, part), silent = TRUE), "try-error") &&
    file.rename(part, cache)
  unlink(part)
  invisible(saved)
}

# the text of this block, so that a separate R process can define the same
# functions. taken from the chunk while the app is being knitted, otherwise
# from the application file in the working folder.
stan_helpers_text <- function() {
  src <- tryCatch(unlist(knitr::knit_code$get("setup-chunk")), error = function(e) NULL)
  if (!any(grepl("^# --- stan helpers: end", src))) {
    for (f in list.files(".", pattern = "[.]Rmd$")) {
      s <- readLines(f, warn = FALSE, encoding = "UTF-8")
      if (any(grepl("^# --- stan helpers: end", s))) { src <- s; break }
    }
  }
  i0 <- grep("^# --- stan helpers: start", src)
  i1 <- grep("^# --- stan helpers: end", src)
  if (length(i0) != 1 || length(i1) != 1 || i1 <= i0) return(NULL)
  paste(src[(i0 + 1):(i1 - 1)], collapse = "\n")
}

# the background process of this R session, if one has been started
stan_job <- function() get0(".epiparam_stan_job", envir = .GlobalEnv, inherits = FALSE)

# while the background process works through its list, the model a user is
# waiting for is written here and compiled next
stan_next_file <- function() {
  file.path(stan_cache_dir(), sprintf("next-%s.txt", Sys.getpid()))
}

# start a separate R process that compiles the models, one at a time. returns
# FALSE when that is not possible (callr missing, or the cache folder cannot be
# written to).
stan_start_build <- function(models) {
  if (!requireNamespace("callr", quietly = TRUE)) return(FALSE)
  if (file.access(stan_cache_dir(), 2) != 0) return(FALSE)
  helpers <- if (exists("STAN_HELPERS_TEXT") && !is.null(STAN_HELPERS_TEXT)) {
    STAN_HELPERS_TEXT
  } else {
    stan_helpers_text()
  }
  if (is.null(helpers)) return(FALSE)
  log <- tempfile("compile-", tmpdir = stan_cache_dir(), fileext = ".log")
  job <- tryCatch(
    callr::r_bg(
      function(helpers, models, cache, next_file, log) {
        Sys.setenv(EPIPARAM_STAN_CACHE = cache, EPIPARAM_STAN_LOG = log)
        eval(parse(text = helpers), envir = globalenv())
        on.exit(unlink(c(next_file, log)), add = TRUE)
        todo <- models
        while (length(todo)) {
          wanted <- tryCatch(readLines(next_file, warn = FALSE)[1],
                             error = function(e) NA, warning = function(w) NA)
          m <- if (!is.na(wanted) && wanted %in% todo) wanted else todo[1]
          todo <- setdiff(todo, m)
          stan_build(m)
        }
      },
      args = list(helpers = helpers, models = models, cache = stan_cache_dir(),
                  next_file = stan_next_file(), log = log),
      # stopped, with the compiler it started, when the app stops
      stdout = log, stderr = "2>&1", supervise = TRUE, cleanup_tree = TRUE
    ),
    error = function(e) NULL
  )
  if (is.null(job)) return(FALSE)
  assign(".epiparam_stan_job", job, envir = .GlobalEnv)
  TRUE
}

# when the app opens: start compiling, in the background, the models this
# computer does not have yet, the default model of each bayesian page first.
# set EPIPARAM_STAN_PREPARE=false to switch this off.
stan_prepare_models <- function(models = c("Lognormal", "Normal", "Weibull", "Gamma")) {
  if (identical(tolower(Sys.getenv("EPIPARAM_STAN_PREPARE")), "false")) return(invisible(FALSE))
  job <- stan_job()
  if (!is.null(job) && job$is_alive()) return(invisible(TRUE))
  have <- file.exists(vapply(models, stan_cache_file, "")) |
    file.exists(vapply(models, stan_bundle_file, ""))
  if (all(have)) return(invisible(FALSE))
  invisible(stan_start_build(models))
}

# progress note, shown in the app while it is running and on the console otherwise
stan_note <- function(msg, console = TRUE) {
  if (requireNamespace("shiny", quietly = TRUE) &&
      !is.null(shiny::getDefaultReactiveDomain())) {
    try(shiny::setProgress(detail = msg), silent = TRUE)
  } else if (console) {
    message(msg)
  }
}

stan_elapsed <- function(secs) {
  secs <- max(0, round(as.numeric(secs)))
  if (secs < 60) paste(secs, "s") else sprintf("%d min %02d s", secs %/% 60, secs %% 60)
}

# wait until the model is compiled, starting the compilation if no one has, and
# keep the user informed meanwhile. returns FALSE when no separate process could
# be used, so the caller compiles in this R session instead.
stan_wait_for_model <- function(model_name) {
  cache   <- stan_cache_file(model_name)
  started <- Sys.time()
  own     <- FALSE
  tick    <- 0
  repeat {
    busy <- stan_lock_busy(model_name)
    if (!busy && file.exists(cache)) return(TRUE)
    # a compilation never takes this long: stop waiting and compile here
    if (difftime(Sys.time(), started, units = "mins") > 20) return(FALSE)
    on_console <- tick %% 15 == 0
    tick <- tick + 1
    job <- stan_job()
    job_alive <- !is.null(job) && job$is_alive()
    if (busy) {
      since <- tryCatch(file.mtime(stan_lock_dir(model_name)), error = function(e) started)
      stan_note(paste0("compiling the ", tolower(model_name), " model for this computer ",
                       "(first run only, usually one to three minutes): ",
                       stan_elapsed(difftime(Sys.time(), since, units = "secs"))), on_console)
    } else if (job_alive) {
      # the background process is busy with another model: this one is next
      try(writeLines(model_name, stan_next_file()), silent = TRUE)
      other <- Filter(stan_lock_busy, setdiff(STAN_MODELS, model_name))
      stan_note(paste0(if (length(other)) paste0("finishing the ", tolower(other[1]), " model; ") else "",
                       "the ", tolower(model_name), " model is compiled next (first run only): ",
                       stan_elapsed(difftime(Sys.time(), started, units = "secs"))), on_console)
    } else if (own) {
      return(TRUE)    # the compilation ended without a model; the caller reports why
    } else {
      if (!stan_start_build(model_name)) return(FALSE)
      own <- TRUE
    }
    Sys.sleep(1)
  }
}

# what to tell the user when the machine has no working C++ toolchain
stan_build_help <- function(model_name, err) {
  paste0(
    "The ", model_name, " model could not be compiled on this computer.\n\n",
    "EpiParam builds the Bayesian models the first time they are used, and that needs\n",
    "a working C++ toolchain:\n",
    "  Windows   install Rtools from https://cran.r-project.org/bin/windows/Rtools/\n",
    "  macOS     run  xcode-select --install  in the Terminal\n",
    "  Linux     install the build tools, for example  sudo apt install build-essential\n\n",
    "Restart R once it is installed, then run the model again.\n\n",
    "The compiler reported:\n", err
  )
}

# the compiled model, from wherever it is available (see the notes above)
get_compiled_stan_model <- function(model_name = "Weibull") {

  code <- get_stan_code(model_name)
  if (is.null(code)) stop("unknown bayesian model: ", model_name, call. = FALSE)

  obj_name <- paste0("stan_model_", model_name)
  keep <- function(mod) {
    assign(obj_name, mod, envir = .GlobalEnv)
    mod
  }

  # already loaded in this session
  if (exists(obj_name, envir = .GlobalEnv)) {
    mod <- get(obj_name, envir = .GlobalEnv)
    if (stan_model_usable(mod, code)) return(mod)
  }

  # compiled on this computer before, or shipped with the app for this installation
  mod <- stan_read_model(stan_cache_file(model_name), code, remove_stale = TRUE)
  if (is.null(mod)) mod <- stan_read_model(stan_bundle_file(model_name), code)
  if (!is.null(mod)) return(keep(mod))

  # otherwise it is compiled here, in a separate R process where possible
  if (stan_wait_for_model(model_name)) {
    mod <- stan_read_model(stan_cache_file(model_name), code)
    if (!is.null(mod)) return(keep(mod))
    err <- tryCatch(paste(readLines(stan_error_file(model_name), warn = FALSE), collapse = "\n"),
                    error = function(e) NULL, warning = function(w) NULL)
    if (!is.null(err)) stop(stan_build_help(model_name, err), call. = FALSE)
  }

  # and in this R session when that is not possible
  stan_note(paste0("compiling the ", tolower(model_name),
                   " model for this computer (first run only, usually one to three minutes)"))
  mod <- tryCatch(
    rstan::stan_model(model_code = code,
                      model_name = paste0("epiparam_", tolower(model_name))),
    error = function(e) stop(stan_build_help(model_name, conditionMessage(e)), call. = FALSE)
  )
  try(saveRDS(mod, stan_cache_file(model_name)), silent = TRUE)
  keep(mod)
}

# --- stan helpers: end --------------------------------------------------

# the block above as text, for the separate R process that compiles the models
STAN_HELPERS_TEXT <- stan_helpers_text()

```

```{r stan-prepare}
# start compiling, in the background, any bayesian model this computer does not
# have yet, so that it is usually ready by the time someone presses run. it
# starts once the page has reached the browser: the page itself is knitted in a
# temporary folder, and the stan_models folder is looked for in the app folder.
# (invisible: onFlushed returns a function, which would otherwise be printed on the page)
if (!is.null(shiny::getDefaultReactiveDomain())) {
  invisible(shiny::onFlushed(stan_prepare_models, once = TRUE))
}
```

```{r font-face}
# the typeface of the app, jost (sil open font license, see fonts/OFL.txt), is
# embedded in the page, so the app looks the same on windows, macos and linux,
# with or without an internet connection. to use another typeface, change
# --epiparam-font in the style sheet at the top of this file.
local({
  font_file <- "fonts/Jost-VariableFont_wght.woff2"
  if (file.exists(font_file)) {
    htmltools::tags$style(paste0(
      "@font-face { font-family: 'Jost'; font-style: normal; font-weight: 100 900; ",
      "font-display: swap; src: url(data:font/woff2;base64,",
      base64enc::base64encode(font_file), ") format('woff2'); }"
    ))
  }
})
```

```{r global-downloads, context="server"}

# global mock data handlers
output$downloadMockIncubation <- downloadHandler(
  filename = function() { paste0("mock_incubation-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("Busan_real.csv"), file) }
)

output$downloadMockserial <- downloadHandler(
  filename = function() { paste0("mock_serial-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("Serial.csv"), file) }
)

output$downloadMockKappa <- downloadHandler(
  filename = function() { paste0("mock_kappa_data-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("kappa_data.csv"), file) }
)

output$downloadMockRt <- downloadHandler(
  filename = function() { paste0("mock_Rt_data-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("rt.csv"), file, row.names = FALSE) }
)

output$downloadMockIncubationBayes <- downloadHandler(
  filename = function() { paste0("mock_incubation_bayesian-", Sys.Date(), ".csv") },
  content = function(file) { 
    write.csv(read.csv("incubation_bayesian.csv"), file, row.names = FALSE) 
  }
)

output$downloadMockSerialBayes <- downloadHandler(
  filename = function() { paste0("mock_serial_bayesian-", Sys.Date(), ".csv") },
  content = function(file) { 
    write.csv(read.csv("serial_bayesian.csv"), file, row.names = FALSE) 
  }
)

output$downloadUserManual <- downloadHandler(
  filename = "User_manual.pdf",
  content = function(file){
    file.copy("User_manual.pdf", file)
  }
)

```

Incubation Period {data-orientation=rows}
==========================================

Sidebar {.sidebar}
--------------------

```{r inc-sidebar}

useShinyjs(rmd = TRUE)

hr()

fileInput("file1_inc", "Choose CSV File",
          multiple = TRUE,
          accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
hr()
selectInput(inputId = "sex_inc", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_group_inc", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")
hr()

selectInput(inputId = "model_inc", label = "Distribution Model",
            choices = c("Lognormal", "Weibull", "Gamma", "ALL"),
            selected = "Lognormal")

# comparison view toggle for all models
conditionalPanel(
  condition = "input.model_inc == 'ALL'",
  radioButtons("view_inc", "Comparison View:",
               choices = c("Side-by-Side", "Overlaid"),
               selected = "Side-by-Side")
)

hr()
valueBox(
  value = tags$h6("Note: The Incubation Period is estimated using Maximum Likelihood Estimation (MLE).",
                  style = "text-align: justify; font-family: var(--epiparam-font); font-size: 20px; color: white;"),
  color = "#FFFFFF")

hr()
valueBox(
  value = downloadButton("downloadMockIncubation", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

hr()
valueBox(
  value = downloadButton("downloadUserManual", "Download User Manual", icon = icon("download")),
  color = "#FFFFFF"
)

renderImage({
  pfad <- "one_logo.png"
  list(
    src = pfad,
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "display: block; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = F)

```

```{r inc-data-processing}

# incubation data processing pipeline

dataset_inc <- eventReactive(list(input$file1_inc, input$sex_inc, input$age_group_inc), {
  req(input$file1_inc)
  
  dataset <- read.csv(input$file1_inc$datapath)
  dataset <- data.frame(lapply(dataset, trimws), stringsAsFactors = FALSE)
  dataset <- clean_names(dataset)
  
  # validation start
  validate(
    need(all(c("date_of_exposure", "date_of_onset") %in% names(dataset)),
         "incorrect dataset: please upload the incubation data containing 'date of exposure' and 'date of onset'.")
  )
  # validation end
  
  if("sex" %in% names(dataset)) {
    dataset <- dataset %>%
      mutate(sex = case_when(
        grepl("^m", sex, ignore.case = TRUE) ~ "Male",
        grepl("^f", sex, ignore.case = TRUE) ~ "Female",
        TRUE ~ sex
      ))
  }
  
  if("age_group" %in% names(dataset)) {
    dataset$age_group <- dataset$age_group %>%
      trimws() %>%              
      gsub("s$", "", ., ignore.case = TRUE) %>%
      gsub("[^0-9]", "", .)
    dataset$age_group <- suppressWarnings(as.numeric(dataset$age_group))
    dataset <- dataset %>% mutate(age_group = case_when(
      age_group > 0 & age_group <= 19 ~ '0 - 19',
      age_group >= 20 & age_group <= 29 ~ '20 - 29',
      age_group >= 30 & age_group <= 39 ~ '30 - 39',
      age_group >= 40 & age_group <= 49 ~ '40 - 49',
      age_group >= 50 & age_group <= 59 ~ '50 - 59',
      TRUE ~ '≥ 60'
    ))
  }
  
  if (input$sex_inc != "All") dataset <- dataset %>% filter(sex == input$sex_inc)
  if (input$age_group_inc != "All") dataset <- dataset %>% filter(age_group == input$age_group_inc)
  
  dataset
})

# every row of the selection, with the dates as the app read them and a note of
# what became of it. the preview, the check panel and the model all read this,
# so what the user is shown is what the model was given.
inc_checked <- reactive({
  req(dataset_inc())
  d <- dataset_inc()
  d$exposure_read   <- parse_date(d$date_of_exposure)
  d$onset_read      <- parse_date(d$date_of_onset)
  d$incubation_days <- as.numeric(d$onset_read - d$exposure_read)
  d$status <- dplyr::case_when(
    is.na(d$exposure_read) & is.na(d$onset_read) ~ "excluded: neither date could be read",
    is.na(d$exposure_read)                       ~ "excluded: exposure date could not be read",
    is.na(d$onset_read)                          ~ "excluded: onset date could not be read",
    d$incubation_days <= 0                       ~ "excluded: onset falls on or before exposure",
    TRUE                                         ~ "used"
  )
  d
})

get_incu_data <- function() {
  d <- inc_checked()
  data <- d$incubation_days[d$status == "used"]
  data <- data[!is.na(data)]
  if(length(data) == 0) return(NULL)
  data
}

# master fitting logic for incubation
fit_results_inc <- reactive({
  data <- get_incu_data()
  if (is.null(data) || length(data) < 3) return(NULL)
  
  models <- if(input$model_inc == "ALL") c("lognormal", "weibull", "gamma") else tolower(input$model_inc)
  
  res <- list()
  for(m in models) {
    # fit logic
    suppressWarnings({
      fit <- fit_parametric_model(data, m, shift_val = 0)
    })
    if(!is.null(fit)) res[[m]] <- fit
  }
  res
})

```

## Row {data-height="180"}

### Sample Size

```{r inc-sample-size}

renderUI({
  # ensure dataset is available
  req(dataset_inc())
  # the number of incubation periods the model is fitted to, as on the other
  # windows (rows that cannot be used are listed under dataset preview)
  ss <- length(get_incu_data())
  # render the sample size text
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: var(--epiparam-font);",
      ss
    )
  )
})

```

### Model Fit Statistics

```{r inc-fit-stats}

renderDT({
  res <- fit_results_inc()
  data <- get_incu_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = paste("Insufficient data points (N=", length(data), "). Need 3+ incubation period data points."))
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data or fit failed")
  } else {
    if (input$model_inc != "ALL") {
      df <- res[[1]]$raw_params_df
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model, 
          AIC = round(x$aic, 2), 
          BIC = round(x$bic, 2)
        )
      }))
      df <- df[order(df$AIC), ] 
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list("csv", "excel", "pdf"),
      pageLength = nrow(df)
    ),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

### Model Parameters

```{r inc-model-params}

renderDT({
  res <- fit_results_inc()
  data <- get_incu_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = "Insufficient Data")
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data")
  } else {
    fmt_moments <- function(x) {
      data.frame(
        parameter = c(paste(x$model, "Mean"), paste(x$model, "Variance"), "95th Percentile"),
        value = round(c(x$moments["mean"], x$moments["var"], x$moments["ci95"]), 3)
      )
    }
    
    if (input$model_inc != "ALL") {
      df <- fmt_moments(res[[1]])
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model,
          `Mean` = round(x$moments["mean"], 3),
          `Variance` = round(x$moments["var"], 3),
          `95th Percentile` = round(x$moments["ci95"], 3),
          check.names = FALSE
        )
      }))
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list("csv", "excel", "pdf"),
      pageLength = nrow(df)
    ),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Distribution Plots

```{r inc-histograms}

renderHighchart({
  data <- get_incu_data()
  res <- fit_results_inc()
  
  if (is.null(data) || length(data) < 3 || is.null(res) || length(res) == 0) {
    return(highchart() %>% hc_title(text="Insufficient Data (N < 3)"))
  }
  
  col_map <- list(lognormal="#00BFFF", weibull="#FF8C00", gamma="#228B22")
  line_col_overlaid <- list(lognormal="#000000", weibull="#FF0000", gamma="#008000") 
  
  # style definitions
  axis_style <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  
  if (input$model_inc == "ALL") {
    
    # calculate global max y for constant scale
    max_y <- 0
    all_plot_data <- list()
    for(m in names(res)) {
      pd <- get_plot_data_inc(data, res[[m]])
      local_max <- max(c(pd$hist$y, pd$curve$y), na.rm = TRUE)
      if(local_max > max_y) max_y <- local_max
      all_plot_data[[m]] <- pd
    }
    max_y <- max_y * 1.05
    
    # sub-case a: overlaid (combined) view
    if (input$view_inc == "Overlaid") {
      pd <- all_plot_data[[1]] # base hist data
      
      hc <- highchart() %>%
        hc_chart(type = "line", style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Combined Model Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"), # fix for axis alignment
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        hc_xAxis(
          title = list(text = "Incubation Periods", style = axis_style), 
          allowDecimals = FALSE, 
          lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = 0, minPadding = 0, startOnTick = TRUE # fix for gap removal
        ) %>%
        hc_yAxis(
          title = list(text = "Density", style = axis_style), 
          gridLineWidth = 0, 
          lineWidth = 1, lineColor="#000000", tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = 0, max = max_y, minPadding = 0, startOnTick = TRUE # fix for gap removal
        ) %>%
        # add histogram
        hc_add_series(
          data = list_parse(pd$hist),
          type = "column",
          name = "Observed Data",
          color = "#D3D3D3", 
          borderColor = "#FFFFFF", borderWidth = 0.25,
          pointRange = 1,
          zIndex = 0
        )
      
      # add curves
      for(m in names(res)) {
        nm <- tolower(m)
        pd_model <- all_plot_data[[m]]
        
        hc <- hc %>% hc_add_series(
          data = pd_model$curve,
          type = "line",
          name = paste(m, "Fit"),
          color = line_col_overlaid[[nm]], 
          lineWidth = 2, marker = list(enabled = FALSE),
          zIndex = 2
        )
      }
      
      hc <- hc %>% hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)))
      return(hc)
      
    } else {
      # sub-case b: side-by-side (composite) view
      hc <- highchart() %>%
        hc_chart(type = "column", style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Distribution Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"), # User Fix for alignment
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        
        # define 3 x-axes
        hc_xAxis_multiples(
          list(title = list(text = "Lognormal", style = axis_style), left = "0%", width = "30%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style = label_style), min=0, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Weibull", style = axis_style), left = "35%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style = label_style), min=0, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Gamma", style = axis_style), left = "70%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style = label_style), min=0, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        # define 3 y-axes
        hc_yAxis_multiples(
          list(title = list(text = "Density", style = axis_style), left = "0%", width = "30%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style = label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style = axis_style), left = "35%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style = label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style = axis_style), left = "70%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style = label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        hc_exporting(
          enabled = TRUE,
          buttons = list(contextButton = list(menuItems = menu_items)),
          chartOptions = list(chart = list(backgroundColor = "#FFFFFF"))
        ) %>%
        hc_legend(enabled = FALSE)
      
      axes_idx <- list(lognormal=0, weibull=1, gamma=2)
      
      for(m in names(res)) {
        nm <- tolower(m)
        idx <- axes_idx[[nm]]
        if(is.null(idx)) next 
        
        pd <- all_plot_data[[m]]
        
        hc <- hc %>%
          hc_add_series(
            data = list_parse(pd$hist),
            type = "column",
            color = col_map[[nm]],
            xAxis = idx, yAxis = idx,
            borderColor = "#FFFFFF", borderWidth = 0.25,
            pointRange = 1, 
            name = paste(m, "Hist")
          ) %>%
          hc_add_series(
            data = pd$curve,
            type = "line",
            color = "#000000",
            xAxis = idx, yAxis = idx,
            lineWidth = 1.5, marker = list(enabled = FALSE),
            name = paste(m, "Fit")
          )
      }
      return(hc)
    }
    
  } else {
    # case 2: single model
    m <- names(res)[[1]]
    pd <- get_plot_data_inc(data, res[[1]])
    nm <- tolower(m)
    
    highchart() %>%
      hc_chart(style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
      hc_title(text = paste(str_to_title(m), "Distribution"), style = list(fontSize="16px", fontWeight="bold")) %>%      
      hc_plotOptions(
        series = list(pointPlacement = "on"), # fix for alignment
        column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
      ) %>%
      hc_add_series(
        data = list_parse(pd$hist),
        type = "column",
        name = "Histogram",
        color = col_map[[nm]], 
        borderColor = "#FFFFFF", borderWidth = 0.25,
        pointRange = 1
      ) %>%
      hc_add_series(
        data = pd$curve,
        type = "line",
        name = "Fit",
        color = "#000000",
        lineWidth = 1.5, marker = list(enabled = FALSE)
      ) %>%
      hc_xAxis(
        title = list(text = "Incubation Periods", style = axis_style), 
        min = floor(min(data))-1, tickInterval = 1,
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = 0, minPadding = 0, startOnTick = TRUE # fix for gap removal
      ) %>%
      hc_yAxis(
        title = list(text = "Density", style = axis_style),
        gridLineWidth = 0,
        lineWidth = 1, lineColor = "#000000", tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = 0, minPadding = 0, startOnTick = TRUE # fix for gap removal
      ) %>%
      hc_exporting(
        enabled = TRUE,
        buttons = list(contextButton = list(menuItems = menu_items)),
        chartOptions = list(chart = list(backgroundColor = "#FFFFFF"))
      )
  }
})

```

### Dataset Preview {.data-preview}

```{r inc-data-check}
renderUI({
  d <- inc_checked()
  data_check_panel(
    n_read = nrow(d),
    n_used = sum(d$status == "used"),
    problems = problems_from_status(d$status, list(
      "excluded: exposure date could not be read" =
        unreadable_values(d$date_of_exposure, d$exposure_read),
      "excluded: onset date could not be read" =
        unreadable_values(d$date_of_onset, d$onset_read),
      "excluded: neither date could be read" =
        c(unreadable_values(d$date_of_exposure, d$exposure_read),
          unreadable_values(d$date_of_onset, d$onset_read))
    ))
  )
})
```

```{r inc-dataset-preview}
renderDataTable({
  # the values as the app read them come first, so a date it could not read is
  # visible without scrolling
  dplyr::relocate(inc_checked(), status, exposure_read, onset_read, incubation_days)
}, options = list(
  pageLength = 10,
  scrollX = TRUE,
  paging = TRUE, 
  striped = TRUE, 
  bordered = TRUE,
  hover = TRUE  
), rownames = FALSE, fillContainer = FALSE)

```


Incubation Period (Bayesian) {data-orientation=rows}
==========================================

Sidebar {.sidebar}
--------------------

```{r inc-bayes-sidebar}

hr()
fileInput("file_bayes_inc", "Choose CSV File",
          multiple = FALSE,
          accept = c("text/csv", "text/comma-separated-values", ".csv"))

hr()
# stratification inputs
selectInput(inputId = "sex_bayes_inc", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_bayes_inc", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")

hr()
# model selection
selectInput("bayes_model_select", "Distribution Model", 
            choices = c("Lognormal","Weibull", "Gamma"), 
            selected = "Lognormal")

sliderInput("iter_count_inc", "Iterations (Max 4000)", value = 1000, min = 500, max = 4000, step = 100)

hr()
actionButton("run_bayes_inc", "Run Bayesian Model", icon = icon("play"), class = "btn-primary", width = "100%")

hr()
valueBox(value = tags$h6("Note: This model uses Bayesian MCMC (rstan). Sampling usually takes under a minute. The first time EpiParam runs on a computer, the models are compiled once, in the background from the moment the app opens (one to three minutes each). The random seed is fixed, so the same data and settings always give the same result.",
                         style = "text-align: justify; font-family: var(--epiparam-font); font-size: 16px; color: white;"), 
         color = "#333333")

hr()

valueBox(
  value = downloadButton("downloadMockIncubationBayes", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

renderImage({
  pfad <- "one_logo.png"
  list(
    src = pfad,
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "display: block; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = F)

```


```{r inc-bayes-data-processing}

# global data loading (runs once per file upload)
bayes_inc_global <- reactive({
  req(input$file_bayes_inc)
  
  # robust read
  tryCatch({
    df <- read.csv(input$file_bayes_inc$datapath, stringsAsFactors = FALSE)
  }, error = function(e) {
    validate(need(FALSE, "error reading file. please ensure it is a valid csv."))
  })
  
  df <- clean_names(df) 
  
  # validation start
  validate(
    need(all(c("symptom_onset", "exp_date_min", "exp_date_max") %in% names(df)),
         "incorrect dataset: please upload incubation bayesian data containing 'symptom onset', 'exp date min', and 'exp date max'.")
  )
  # validation end
  
  # standardize sex
  if("sex" %in% names(df)) {
    df <- df %>% mutate(sex = case_when(
        grepl("^m", sex, ignore.case = TRUE) ~ "Male",
        grepl("^f", sex, ignore.case = TRUE) ~ "Female",
        TRUE ~ sex
      ))
  }
  
  # standardize age group
  if("age" %in% names(df)) {
    df$age <- suppressWarnings(as.numeric(df$age))
    df <- df %>% mutate(age_group = case_when(
      age > 0 & age <= 19 ~ '0 - 19',
      age >= 20 & age <= 29 ~ '20 - 29',
      age >= 30 & age <= 39 ~ '30 - 39',
      age >= 40 & age <= 49 ~ '40 - 49',
      age >= 50 & age <= 59 ~ '50 - 59',
      TRUE ~ '≥ 60'
    ))
  }
  
  # parse dates globally
  df <- df %>%
    mutate(
      date_symp = parse_date(symptom_onset),
      date_min  = parse_date(exp_date_min),
      date_max  = parse_date(exp_date_max)
    )
  
  df
})

# filtering logic
bayes_inc_data <- reactive({
  req(bayes_inc_global())
  df <- bayes_inc_global()
  
  # filter raw data
  if (input$sex_bayes_inc != "All" && "sex" %in% names(df)) {
    df <- df %>% filter(sex == input$sex_bayes_inc)
  }
  if (input$age_bayes_inc != "All" && "age_group" %in% names(df)) {
    df <- df %>% filter(age_group == input$age_bayes_inc)
  }
  
  # capture count before cleaning
  n_before_cleaning <- nrow(df)
  
  # model prep logic
  df <- df %>%
    filter(!is.na(date_symp)) %>% # must have symptom date
    mutate(
      tSymptomOnset = as.integer(date_symp - as.Date("2019-12-31")),
      tStartExposure = as.integer(date_min - as.Date("2019-12-31")),
      tEndExposure = as.integer(date_max - as.Date("2019-12-31"))
    ) %>%
    # logic update!!
    # keep rows with at least one exposure date
    filter(!is.na(tStartExposure) | !is.na(tEndExposure)) %>%
    # impute missing end date (assume exposure continued until symptoms)
    mutate(tEndExposure = ifelse(is.na(tEndExposure), tSymptomOnset, tEndExposure)) %>%
    # impute missing start date (assume exposure started 14 days before end date)
    mutate(tStartExposure = ifelse(is.na(tStartExposure), tEndExposure, tStartExposure)) %>% # you can update the window here
    # final safety check: end exposure cannot be after symptoms
    mutate(tEndExposure = ifelse(tEndExposure > tSymptomOnset, tSymptomOnset, tEndExposure)) %>%
    # ensure start is before end
    filter(tStartExposure <= tEndExposure)
  
  # capture count after
  n_final <- nrow(df)
  n_dropped <- n_before_cleaning - n_final
  
  # notify user
  if (n_dropped > 0) {
    showNotification(
      paste0("data loaded: ", n_final, " valid rows used. (", n_dropped, " rows removed: missing both exposure dates or invalid timeline)"),
      type = "warning",
      duration = 5 
    )
  } else {
    showNotification(
      paste0("data successfully loaded: ", n_final, " rows ready."),
      type = "message",
      duration = 3 
    )
  }
  
  df
})

# every row of the selection, with the dates as the app read them, the exposure
# window it ended up using and a note of what became of the row. this repeats
# the preparation the model does, so the preview shows the model's own view of
# the file rather than the raw text.
bayes_inc_checked <- reactive({
  req(bayes_inc_global())
  d <- bayes_inc_global()
  if (input$sex_bayes_inc != "All" && "sex" %in% names(d)) {
    d <- d %>% filter(sex == input$sex_bayes_inc)
  }
  if (input$age_bayes_inc != "All" && "age_group" %in% names(d)) {
    d <- d %>% filter(age_group == input$age_bayes_inc)
  }
  
  origin <- as.Date("2019-12-31")
  tS <- as.integer(d$date_symp - origin)
  tB <- as.integer(d$date_min  - origin)
  tE <- as.integer(d$date_max  - origin)
  tE <- ifelse(is.na(tE), tS, tE)          # no end date: exposure ran to onset
  tB <- ifelse(is.na(tB), tE, tB)          # no start date: window has no width
  tE <- ifelse(tE > tS, tS, tE)            # exposure cannot end after onset
  
  d$onset_read           <- d$date_symp
  d$exposure_from        <- d$date_min
  d$exposure_to          <- d$date_max
  d$exposure_window_days <- tE - tB
  d$days_to_onset        <- tS - tE
  d$status <- dplyr::case_when(
    is.na(d$date_symp)                    ~ "excluded: symptom onset date could not be read",
    is.na(d$date_min) & is.na(d$date_max) ~ "excluded: no exposure date given",
    tB > tE                               ~ "excluded: exposure window ends before it starts",
    TRUE                                  ~ "used"
  )
  d
})

# modeling
bayes_inc_fit <- eventReactive(input$run_bayes_inc, {
  
  # lock button
  shinyjs::disable("run_bayes_inc")
  
  # unlock button when done
  on.exit({
    shinyjs::enable("run_bayes_inc")
  })
  
  # double clean
  gc()
  gc() 
  
  req(bayes_inc_data())
  df <- bayes_inc_data()

  validate(
    need(nrow(df) >= 5, "insufficient data: you need at least 5 observations to run this model.")
  )

  stan_data <- list(
    N = nrow(df),
    tStartExposure = df$tStartExposure,
    tEndExposure = df$tEndExposure,
    tSymptomOnset = df$tSymptomOnset
  )

  withProgress(message = 'running bayesian model...', detail = 'initializing...', value = 0, {

    mod <- get_compiled_stan_model(input$bayes_model_select)
    incProgress(0.1, detail = "sampling chains...")

    fit <- rstan::sampling(
      object = mod,
      data = stan_data,
      chains = 4,
      cores = min(4, parallel::detectCores(), na.rm = TRUE),
      seed = STAN_SEED,
      iter = input$iter_count_inc,
      warmup = input$iter_count_inc / 2,
      refresh = 0,
      pars = c("alphaInc", "sigmaInc"), # keep light!
      control = list(adapt_delta = 0.95)
    )

    incProgress(0.8, detail = "extracting posteriors...")

    params <- rstan::extract(fit)

    list(
      fit_obj = fit,
      iterations = input$iter_count_inc,
      model_name = input$bayes_model_select,
      alpha = params$alphaInc,
      sigma = params$sigmaInc,
      alpha_summary = quantile(params$alphaInc, probs = c(0.025, 0.5, 0.975)),
      sigma_summary = quantile(params$sigmaInc, probs = c(0.025, 0.5, 0.975))
    )
  })
})

```

## Row {data-height="180"}

### Sample Size

```{r inc-bayes-sample-size}

renderUI({
  req(bayes_inc_data())
  
  ss <- nrow(bayes_inc_data())
  
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 40px; font-weight: bold; margin-bottom: 5px; text-align: center; font-family: var(--epiparam-font); color: #000000;",
      ss
    ),
    htmltools::div(
      style = "font-size: 14px; text-align: center; color: #666;",
      "Observations"
    )
  )
})

```


### Model Statistics (Median & 95% CrI)

```{r inc-bayes-model-stats}

renderDT({
  req(bayes_inc_fit())
  res <- bayes_inc_fit()
  
  # dynamic parameter labels & mean calculation
  if (res$model_name == "Lognormal") {
    p_names <- c("MeanLog (Alpha)", "SDLog (Sigma)")
    # mean of lognormal = exp(mu + sigma^2/2)
    post_means <- exp(res$alpha + (res$sigma^2)/2)
  } else if (res$model_name == "Gamma") {
    p_names <- c("Shape (Alpha)", "Rate (Sigma)")
    # mean of gamma = shape / rate
    post_means <- res$alpha / res$sigma
  } else {
    # weibull
    p_names <- c("Shape (Alpha)", "Scale (Sigma)")
    # mean of weibull = scale * gamma(1 + 1/shape)
    post_means <- res$sigma * gamma(1 + 1/res$alpha)
  }
  
  df_stats <- data.frame(
    Parameter = p_names,
    `Median` = round(c(res$alpha_summary[2], res$sigma_summary[2]), 3),
    `Lower_95_CrI` = round(c(res$alpha_summary[1], res$sigma_summary[1]), 3),
    `Upper_95_CrI` = round(c(res$alpha_summary[3], res$sigma_summary[3]), 3)
  )
  
  mean_summ <- quantile(post_means, probs = c(0.025, 0.5, 0.975))
  
  df_mean <- data.frame(
    Parameter = "Mean Incubation Period",
    `Median` = round(mean_summ[2], 3),
    `Lower_95_CrI` = round(mean_summ[1], 3),
    `Upper_95_CrI` = round(mean_summ[3], 3)
  )
  
  final_df <- rbind(df_stats, df_mean)
  
  fname <- paste0("Incubation_Bayesian_Stats_", Sys.Date())
  
  datatable(
    final_df, 
    rownames = FALSE, 
    extensions = 'Buttons',
    options = list(
      dom = 'Brt',
      buttons = list(
        list(extend = "csv", filename = fname),
        list(extend = "excel", filename = fname),
        list(extend = "pdf", filename = fname)
      ),
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
  ) %>%
    formatStyle(columns = "Parameter", fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------


### Dataset Preview {.data-preview}

```{r inc-bayes-dataset-preview}

renderUI({
  d <- bayes_inc_checked()
  data_check_panel(
    n_read = nrow(d),
    n_used = sum(d$status == "used"),
    problems = c(
      problems_from_status(d$status, list(
        "excluded: symptom onset date could not be read" =
          unreadable_values(d$symptom_onset, d$date_symp)
      )),
      if (any(d$status == "used" & d$exposure_window_days == 0))
        list(sprintf(paste("%d of the rows used have no exposure window: only one exposure date",
                           "was given, so the app treats infection as having happened on that day."),
                     sum(d$status == "used" & d$exposure_window_days == 0)))
    )
  )
})

renderDataTable({
  # the model's view of the file: parsed dates, the window it used, and status
  dplyr::relocate(bayes_inc_checked(), status, onset_read, exposure_from,
                  exposure_to, exposure_window_days, days_to_onset)
}, options = list(
  pageLength = 10,
  scrollX = TRUE,
  paging = TRUE, 
  striped = TRUE, 
  bordered = TRUE, 
  hover = TRUE  
), rownames = FALSE, fillContainer = FALSE)

```


### Posterior Distribution (Median Fit)

```{r inc-bayes-posterior-dist}

renderHighchart({
  req(bayes_inc_fit())
  res <- bayes_inc_fit()
  
  alpha_est <- median(res$alpha)
  sigma_est <- median(res$sigma)
  x_vals <- seq(0, 30, length.out = 200)
  
  # dynamic distribution logic
  if (res$model_name == "Gamma") {
    y_vals <- dgamma(x_vals, shape = alpha_est, rate = sigma_est)
  } else if (res$model_name == "Lognormal") {
    y_vals <- dlnorm(x_vals, meanlog = alpha_est, sdlog = sigma_est)
  } else {
    y_vals <- dweibull(x_vals, shape = alpha_est, scale = sigma_est)
  }
  
  model_label <- tools::toTitleCase(res$model_name)
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  
  # theme consistency
  axis_title_style <- list(color = "#000000", fontSize = "16px", fontWeight = "bold", fontFamily = "Jost")
  axis_label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold", fontFamily = "Jost")
  
  highchart() %>%
    hc_chart(style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
    
    hc_title(
      text = paste0("Posterior Predictive Distribution (", model_label, ")"),
      style = list(fontSize = "18px", fontWeight = "bold", fontFamily = "Jost", color = "#000000")
    ) %>%
    
    hc_subtitle(
      text = paste0("Based on ", res$iterations, " MCMC iterations"),
      style = list(fontSize = "14px", fontFamily = "Jost", color = "#666666")
    ) %>%
    
    hc_exporting(
      enabled = TRUE, 
      filename = paste0("Incubation_Posterior_Distribution_", Sys.Date()),
      buttons = list(contextButton = list(menuItems = menu_items))
    ) %>%
    
    hc_add_series(
      data = data.frame(x = x_vals, y = y_vals),
      type = "area",
      name = "Posterior Density",
      color = "rgba(65, 105, 225, 0.5)",
      marker = list(enabled = FALSE),      
      lineWidth = 2
    ) %>%
    
    # axis styling
    hc_xAxis(
      title = list(text = "Incubation Period (Days)", style = axis_title_style),
      labels = list(style = axis_label_style),
      min = 0, minPadding = 0, startOnTick = TRUE,
      lineColor = "#000000", lineWidth = 1, tickColor = "#000000"
    ) %>%
    
    hc_yAxis(
      title = list(text = "Density", style = axis_title_style),
      labels = list(style = axis_label_style),
      min = 0, minPadding = 0, startOnTick = TRUE,
      gridLineWidth = 0, 
      lineColor = "#000000", lineWidth = 1, tickColor = "#000000"
    ) %>%
    
    hc_tooltip(valueDecimals = 4, style = list(fontFamily = "Jost"))
})

```

### Trace Plot (Convergence Check)

```{r inc-bayes-trace-plot}

renderPlotly({
  req(bayes_inc_fit())
  
  # base plot with grid lines but customized fonts and black axis
  p <- rstan::traceplot(bayes_inc_fit()$fit_obj, 
                        pars = c("alphaInc", "sigmaInc"),
                        linewidth = 0.25) + 
    theme_bw() + 
    ggtitle("Trace Plot: Alpha & Sigma") +
    theme(
      text = element_text(family = "Jost"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10, color = "black"),
      axis.line = element_line(color = "black"), 
      strip.background = element_rect(fill = "#f0f0f0"), 
      strip.text = element_text(face = "bold")
    )
  
  ggplotly(p) %>%
    config(
      toImageButtonOptions = list(
        format = "png",
        filename = paste0("Incubation_Trace_Plot_", Sys.Date()),
        height = 500,
        width = 700,
        scale = 2 
      )
    ) %>%
    layout(font = list(family = "Jost"))
})

```

Serial Interval {data-orientation=rows}
=========================================

Sidebar {.sidebar}
---------------------

```{r si-sidebar}
hr()
fileInput("file1_si", "Choose CSV File",
          multiple = TRUE,
          accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
hr()
selectInput(inputId = "sex_si", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_group_si", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")
hr()

selectInput("model_si", "Distribution Model:",
            choices = c("Normal", "Weibull", "Gamma", "Lognormal", "ALL"),
            selected = "Normal")

# comparison view toggle for si
conditionalPanel(
  condition = "input.model_si == 'ALL'",
  radioButtons("view_si", "Comparison View:",
               choices = c("Side-by-Side", "Overlaid"),
               selected = "Side-by-Side")
)

hr()

valueBox(
  value = downloadButton("downloadMockserial", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

hr()

valueBox(
  value = tags$h6("Note: The Weibull, Gamma and Lognormal models are defined only for positive values, so they are fit on the serial intervals moved to the right by k days, and the estimates are moved back. k is taken from the data, so negative serial intervals are kept rather than dropped. The Normal model is fit on the raw values.",
                  style = "text-align: justify; font-family: var(--epiparam-font); font-size: 16px; color: white;"),
  color = "#FFFFFF")

hr()
# logo image
renderImage({
  list(
    src = "one_logo.png",
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "display: block; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = FALSE)

```


```{r si-data-processing}

# data processing for serial interval
dataset_si <- eventReactive(list(input$file1_si, input$sex_si, input$age_group_si), {
  req(input$file1_si)
  
  dataset <- read.csv(input$file1_si$datapath)
  dataset <- data.frame(lapply(dataset, trimws), stringsAsFactors = FALSE)
  dataset <- clean_names(dataset)
  
  # validation start
  validate(
    need(all(c("id", "source_of_infection", "date_of_onset") %in% names(dataset)),
         "incorrect dataset: please upload serial interval data containing 'id', 'source of infection', and 'date of onset'.")
  )
  # validation end
  
  # more robust cleaning
  if("age_group" %in% names(dataset)) {
    dataset$age_group <- dataset$age_group %>%
      gsub("s$", "", ., ignore.case = TRUE) %>%
      gsub("[^0-9]", "", .)
    dataset$age_group <- suppressWarnings(as.numeric(dataset$age_group))
    dataset <- dataset %>% mutate(age_group = case_when(
      age_group > 0 & age_group <= 19 ~ '0 - 19',
      age_group >= 20 & age_group <= 29 ~ '20 - 29',
      age_group >= 30 & age_group <= 39 ~ '30 - 39',
      age_group >= 40 & age_group <= 49 ~ '40 - 49',
      age_group >= 50 & age_group <= 59 ~ '50 - 59',
      TRUE ~ '≥ 60'
    ))
  }
  
  if("sex" %in% names(dataset)) {
    dataset <- dataset %>%
      mutate(sex = case_when(
        grepl("^m", sex, ignore.case = TRUE) ~ "Male",
        grepl("^f", sex, ignore.case = TRUE) ~ "Female",
        TRUE ~ sex
      ))
  }
  
  # keep what was typed, so the check panel can quote the values it could not use
  dataset$id_as_given     <- as.character(dataset$id)
  dataset$source_as_given <- as.character(dataset$source_of_infection)
  dataset$onset_as_given  <- as.character(dataset$date_of_onset)
  
  dataset$id <- as.numeric(gsub("[^0-9]", "", dataset$id))
  dataset$source_of_infection <- as.numeric(gsub("[^0-9]", "", dataset$source_of_infection))
  dataset$date_of_exposure <- parse_date(dataset$date_of_exposure)
  dataset$date_of_onset <- parse_date(dataset$date_of_onset)
  dataset$source_date <- dataset$date_of_onset[match(dataset$source_of_infection, dataset$id)]
  
  if (input$sex_si != "All") dataset <- dataset %>% filter(sex == input$sex_si)
  if (input$age_group_si != "All") dataset <- dataset %>% filter(age_group == input$age_group_si)
  
  dataset
})

# every row of the selection, with the serial interval the app worked out and a
# note of what became of it. a transmission link that points at a case which is
# not in the file is now reported instead of quietly disappearing.
si_checked <- reactive({
  req(dataset_si())
  d <- dataset_si()
  
  required_cols <- c("id", "source_of_infection", "date_of_onset", "source_date")
  if (!all(required_cols %in% names(d))) stop("required columns not found in dataset")
  
  d$serial_days <- as.numeric(d$date_of_onset - d$source_date)
  known_source  <- !is.na(d$source_of_infection) & d$source_of_infection %in% d$id
  d$status <- dplyr::case_when(
    is.na(d$date_of_onset)       ~ "excluded: onset date could not be read",
    is.na(d$source_of_infection) ~ "excluded: no source of infection given",
    !known_source                ~ "excluded: the source of infection is not one of the case ids in this file",
    is.na(d$source_date)         ~ "excluded: the source case has no readable onset date",
    TRUE                         ~ "used"
  )
  d
})

get_serial_data <- function() {
  d <- si_checked()
  v <- d$serial_days[d$status == "used"]
  v <- v[!is.na(v)]
  if (!length(v)) return(NULL)
  v
}

# master fitting logic for si
fit_results_si <- reactive({
  data <- get_serial_data()
  if (is.null(data) || length(data) < 3) return(NULL)
  
  models <- if(input$model_si == "ALL") c("normal", "weibull", "gamma", "lognormal") else tolower(input$model_si)
  
  res <- list()
  for(m in models) {
    # shift logic. weibull, gamma and lognormal are defined only for positive
    # values, so the serial intervals are moved to the right before fitting and
    # the fitted distribution is moved back afterwards, in fit_parametric_model.
    # the shift is taken from the data, k = 1 - min(si), so a negative serial
    # interval is kept instead of being dropped. with nothing negative k is 1,
    # which is what the app did before.
    shift_val <- if(m == "normal") 0 else max(1, 1 - min(data))
    
    # every value is positive after that shift; kept as a guard
    if(m != "normal") {
      valid_indices <- (data + shift_val) > 0
      data_for_fit <- data[valid_indices]
    } else {
      data_for_fit <- data
    }
    
    # check if we still have enough data after filtering
    if(length(data_for_fit) < 3) next
    
    suppressWarnings({
      fit <- fit_parametric_model(data_for_fit, m, shift_val = shift_val)
    })
    
    if(!is.null(fit)) res[[m]] <- fit
  }
  res
})

```


## Row {data-height="180"}

### Sample Size

```{r si-sample-size}

renderUI({
  req(dataset_si())
  # the number of serial intervals the model is fitted to, as on the other
  # windows (rows that cannot be used are listed under dataset preview)
  ss <- length(get_serial_data())
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: var(--epiparam-font);",
      ss
    )
  )
})

```

### Model Fit Statistics

```{r si-model-fit-stats}

renderDT({
  res <- fit_results_si()
  data <- get_serial_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = "Insufficient Data (N < 3).")
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data or fit failed")
  } else {
    if (input$model_si != "ALL") {
      df <- res[[1]]$raw_params_df
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model, 
          AIC = round(x$aic, 2), 
          BIC = round(x$bic, 2)
        )
      }))
      df <- df[order(df$AIC), ]
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(dom = "Brt", buttons = list("csv", "excel", "pdf"), pageLength = nrow(df)),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

### Model Parameters

```{r si-model-params}

renderDT({
  res <- fit_results_si()
  data <- get_serial_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = "Insufficient Data")
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data")
  } else {
    fmt_moments <- function(x) {
      data.frame(
        parameter = c(paste(x$model, "Mean"), paste(x$model, "Variance"), "95th Percentile"),
        value = round(c(x$moments["mean"], x$moments["var"], x$moments["ci95"]), 3)
      )
    }
    
    if (input$model_si != "ALL") {
      df <- fmt_moments(res[[1]])
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model,
          `Mean` = round(x$moments["mean"], 3),
          `Variance` = round(x$moments["var"], 3),
          `95th Percentile` = round(x$moments["ci95"], 3),
          check.names = FALSE
        )
      }))
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(dom = "Brt", buttons = list("csv", "excel", "pdf"), pageLength = nrow(df)),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Distribution Plots

```{r si-histograms}

renderHighchart({
  data <- get_serial_data()
  res <- fit_results_si()
  
  if (is.null(data) || length(data) < 3 || is.null(res) || length(res) == 0) {
    return(highchart() %>% hc_title(text="Insufficient Data (N < 3)"))
  }
  
  # set model colors
  col_map <- list(
    normal = "#DAA520",    # gold
    lognormal = "#00BFFF", # blue
    weibull = "#FF8C00",   # orange
    gamma = "#228B22"      # green
  )
  line_col_overlaid <- list(normal="#0000FF", lognormal="#000000", weibull="#FF0000", gamma="#008000")
  
  # style definitions
  axis_style <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  
  # ensure the first bar isn't cut in half (start x-axis slightly before min integer)
  min_x_val <- floor(min(data)) - 0.5
  
  # case 1: all models selected
  if (input$model_si == "ALL") {
    
    # calculate global max y
    max_y <- 0
    all_plot_data <- list()
    for(m in names(res)) {
      pd <- get_plot_data_si(data, res[[m]])
      local_max <- max(c(pd$hist$y, pd$curve$y), na.rm = TRUE)
      if(local_max > max_y) max_y <- local_max
      all_plot_data[[m]] <- pd
    }
    max_y <- max_y * 1.05
    
    # sub-case a: overlaid (combined) view
    if (input$view_si == "Overlaid") {
      
      pd <- all_plot_data[[1]] # base hist
      
      hc <- highchart() %>%
        hc_chart(type = "line", style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Combined Model Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"), 
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        hc_xAxis(
          title = list(text = "Serial Interval (Days)", style = axis_style), 
          allowDecimals = FALSE, 
          lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = min_x_val, # Updated start point
          minPadding = 0, startOnTick = TRUE
        ) %>%
        hc_yAxis(
          title = list(text = "Density", style = axis_style), 
          gridLineWidth = 0, 
          lineWidth = 1, lineColor="#000000", tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = 0, max = max_y, 
          minPadding = 0, startOnTick = TRUE
        ) %>%
        hc_add_series(
          data = list_parse(pd$hist), type = "column", name = "Observed Data",
          color = "#D3D3D3", borderColor = "#FFFFFF", borderWidth = 0.25,
          pointRange = 2, zIndex = 0
        )
      
      for(m in names(res)) {
        nm <- tolower(m)
        pd_model <- all_plot_data[[m]]
        hc <- hc %>% hc_add_series(
          data = pd_model$curve, type = "line", name = paste(m, "Fit"),
          color = line_col_overlaid[[nm]], lineWidth = 2, marker = list(enabled = FALSE), zIndex = 2
        )
      }
      hc <- hc %>% hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)))
      return(hc)
      
    } else {
      # sub-case b: side-by-side (composite) view with constant scale
      hc <- highchart() %>%
        hc_chart(type = "column", style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Distribution Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"),
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        
        # 4 x-axes
        hc_xAxis_multiples(
          list(title = list(text = "Normal", style=axis_style), left = "0%", width = "22%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Weibull", style=axis_style), left = "26%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Gamma", style=axis_style), left = "52%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Lognormal", style=axis_style), left = "78%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        # 4 y-axes (aligned, same scale)
        hc_yAxis_multiples(
          list(title = list(text = "Density", style=axis_style), left = "0%", width = "22%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style=axis_style), left = "26%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style=axis_style), left = "52%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style=axis_style), left = "78%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)), chartOptions = list(chart = list(backgroundColor = "#FFFFFF"))) %>%
        hc_legend(enabled = FALSE)
      
      axes_idx <- list(normal=0, weibull=1, gamma=2, lognormal=3)
      for(m in names(res)) {
        nm <- tolower(m)
        idx <- axes_idx[[nm]]
        if(is.null(idx)) next 
        
        pd <- all_plot_data[[m]]
        
        hc <- hc %>%
          hc_add_series(
            data = list_parse(pd$hist), type = "column", color = col_map[[nm]],
            xAxis = idx, yAxis = idx, 
            borderColor = "#FFFFFF", borderWidth = 0.25,
            pointRange = 2, 
            name = paste(m, "Hist")
          ) %>%
          hc_add_series(
            data = pd$curve, type = "line", color = "#000000",
            xAxis = idx, yAxis = idx, 
            lineWidth = 1.5, marker = list(enabled = FALSE),
            name = paste(m, "Fit")
          )
      }
      return(hc)
    }
    
  } else {
    # case 2: single model plot
    m <- names(res)[[1]]
    pd <- get_plot_data_si(data, res[[1]])
    nm <- tolower(m)
    
    highchart() %>%
      hc_chart(style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
      hc_title(text = paste(str_to_title(m), "Distribution"), style = list(fontSize="16px", fontWeight="bold")) %>%      
      hc_plotOptions(
        series = list(pointPlacement = "on"), 
        column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
      ) %>%
      hc_add_series(
        data = list_parse(pd$hist), type = "column", name = "Histogram",
        color = col_map[[nm]], borderColor = "#FFFFFF", borderWidth = 0.25,
        pointRange = 2
      ) %>%
      hc_add_series(
        data = pd$curve, type = "line", name = "Fit",
        color = "#000000", lineWidth = 1.5, marker = list(enabled = FALSE)
      ) %>%
      hc_xAxis(
        title = list(text = "Serial interval (days)", style = axis_style), 
        min = min_x_val, tickInterval = 2,
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        minPadding = 0, startOnTick = TRUE
      ) %>%
      hc_yAxis(
        title = list(text = "Density", style = axis_style),
        gridLineWidth = 0, lineWidth = 1, lineColor = "#000000", tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = 0, minPadding = 0, startOnTick = TRUE
      ) %>%
      hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)), chartOptions = list(chart = list(backgroundColor = "#FFFFFF")))
  }
})

```

### Dataset Preview {.data-preview}

```{r}

renderUI({
  d <- si_checked()
  known_source <- !is.na(d$source_of_infection) & d$source_of_infection %in% d$id
  data_check_panel(
    n_read = nrow(d),
    n_used = sum(d$status == "used"),
    problems = problems_from_status(d$status, list(
      "excluded: onset date could not be read" =
        unreadable_values(d$onset_as_given, d$date_of_onset),
      "excluded: the source of infection is not one of the case ids in this file" =
        unique(d$source_as_given[!known_source & !is.na(d$source_of_infection)])
    ))
  )
})

renderDataTable({
  d <- si_checked()
  
  # ensure date columns are formatted as strings in yyyy-mm-dd
  d$date_of_exposure <- as.Date(d$date_of_exposure)
  d$date_of_onset <- as.Date(d$date_of_onset)
  d$source_date <- as.Date(d$source_date)
  
  # what the app made of the file comes first
  dataset <- dplyr::relocate(d, status, id, source_of_infection,
                             date_of_onset, source_date, serial_days)
  
  datatable(
    dataset,
    options = list(
      pageLength = 10,
      scrollX = TRUE,
      paging = TRUE,
      striped = TRUE,
      bordered = TRUE,
      hover = TRUE
    ),
    rownames = FALSE,
    fillContainer = FALSE
  )
})

```

Serial Interval (Bayesian) {data-orientation=rows}
==========================================

Sidebar {.sidebar}
--------------------

```{r si-bayes-sidebar}

hr()
fileInput("file_bayes_si", "Choose CSV File",
          multiple = FALSE,
          accept = c("text/csv", "text/comma-separated-values", ".csv"))

hr()
# stratification inputs
selectInput(inputId = "sex_bayes_si", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_bayes_si", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")

hr()
# model selection
selectInput("bayes_model_select_si", "Distribution Model", 
            choices = c("Normal","Weibull", "Gamma", "Lognormal"), 
            selected = "Normal")

sliderInput("iter_count_si", "Iterations (Max 4000)", value = 1000, min = 500, max = 4000, step = 100)

hr()
actionButton("run_bayes_si", "Run Bayesian Model", icon = icon("play"), class = "btn-primary", width = "100%")

hr()
valueBox(value = tags$h6("Note: This model estimates Serial Interval by linking Infectors to Infectees. Ensure 'Source of infection' IDs match 'id'. The random seed is fixed, so the same data and settings always give the same result.",
                         style = "text-align: justify; font-family: var(--epiparam-font); font-size: 16px; color: white;"), 
         color = "#333333")

hr()

valueBox(
  value = downloadButton("downloadMockSerialBayes", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

renderImage({
  pfad <- "one_logo.png"
  list(
    src = pfad,
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "display: block; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = F)

```


```{r si-bayes-data-processing}

# data loading/processing
bayes_si_global <- reactive({
  req(input$file_bayes_si)
  tryCatch({
    df <- read.csv(input$file_bayes_si$datapath, stringsAsFactors = FALSE)
  }, error = function(e) {
    validate(need(FALSE, "error reading file. please ensure it is a valid csv."))
  })
  names(df) <- tolower(names(df))
  names(df) <- gsub("\\.", "_", names(df))
  
  # validation start
  validate(
    need(all(c("id", "source_of_infection", "date_of_onsetmin", "date_of_onsetmax") %in% names(df)),
         "incorrect dataset: please upload serial bayesian data containing 'id', 'source of infection', 'date of onsetmin', and 'date of onsetmax'.")
  )
  # validation end
  
  # standardize sex
  if("sex" %in% names(df)) {
    df <- df %>% mutate(sex = case_when(
      grepl("^m", sex, ignore.case = TRUE) ~ "Male",
      grepl("^f", sex, ignore.case = TRUE) ~ "Female",
      TRUE ~ sex
    ))
  }
  
  # standardize age group
  col_age <- grep("age.*group", names(df), value = TRUE)[1]
  if(!is.na(col_age)) {
    df$age_clean <- df[[col_age]]
    df <- df %>% mutate(age_group = case_when(
      grepl("0-19|under 10|10s", age_clean, ignore.case=T) ~ '0 - 19',
      grepl("20s|20-29", age_clean, ignore.case=T) ~ '20 - 29',
      grepl("30s|30-39", age_clean, ignore.case=T) ~ '30 - 39',
      grepl("40s|40-49", age_clean, ignore.case=T) ~ '40 - 49',
      grepl("50s|50-59", age_clean, ignore.case=T) ~ '50 - 59',
      grepl("60s|70s|80s|60", age_clean, ignore.case=T) ~ '≥ 60',
      TRUE ~ 'Unknown'
    ))
  }
  
  df$id_clean <- trimws(gsub("#", "", df$id))
  df$source_clean <- trimws(gsub("#", "", df$source_of_infection))
  df$onset_min <- parse_date(df$date_of_onsetmin)
  df$onset_max <- parse_date(df$date_of_onsetmax)
  
  df
})

# filtering logic 
bayes_si_data <- reactive({
  req(bayes_si_global())
  
  # clean memory
  gc() 
  
  df <- bayes_si_global()
  
  df_infectors <- df 
  df_infectees <- df
  
  if (input$sex_bayes_si != "All" && "sex" %in% names(df)) {
    df_infectees <- df_infectees %>% filter(sex == input$sex_bayes_si)
  }
  if (input$age_bayes_si != "All" && "age_group" %in% names(df)) {
    df_infectees <- df_infectees %>% filter(age_group == input$age_bayes_si)
  }
  
  n_before_cleaning <- nrow(df_infectees)
  
  df_merged <- df_infectees %>%
    inner_join(df_infectors %>% dplyr::select(id_clean, onset_min, onset_max), 
               by = c("source_clean" = "id_clean"), 
               suffix = c("_infectee", "_infector"))
  
  df_final <- df_merged %>%
    filter(!is.na(onset_min_infectee), !is.na(onset_min_infector)) %>%
    mutate(
      tStartExposure = as.integer(onset_min_infector - as.Date("2019-12-31")),
      tEndExposure = as.integer(onset_max_infector - as.Date("2019-12-31")),
      tSymptomOnset = as.integer(onset_min_infectee - as.Date("2019-12-31")) + 
        as.integer(onset_max_infectee - onset_min_infectee)/2
    ) 
  
  # weibull, gamma and lognormal need a positive serial interval. rather than
  # dropping the pairs where the infectee fell ill before, or on the same day
  # as, the infector, the onsets are moved to the right by k days and the
  # fitted distribution is moved back when the results are shown. k covers the
  # whole exposure window, so every draw of tE stays below the onset. when no
  # pair needs it, k is 0 and nothing about the fit changes.
  si_shift <- 0
  if (input$bayes_model_select_si != "Normal" && nrow(df_final) > 0) {
    need <- max(df_final$tEndExposure - df_final$tSymptomOnset)
    if (need >= 0) {
      si_shift <- need + 1
      df_final$tSymptomOnset <- df_final$tSymptomOnset + si_shift
    }
  }
  df_final$si_shift <- si_shift
  
  n_final <- nrow(df_final)
  n_dropped <- n_before_cleaning - n_final
  
  if (n_dropped > 0) {
    showNotification(
      paste0("data loaded: ", n_final, " linked pairs found. (", n_dropped, " rows removed: unlinked or invalid dates)"),
      type = "warning", 
      duration = 5
    )
  } else {
    showNotification(
      paste0("data successfully loaded: ", n_final, " pairs ready."),
      type = "message", 
      duration = 3
    )
  }
  
  df_final
})

# every row of the selection, with the link the app resolved and a note of what
# became of it. a source of infection that is not itself a case in the file is
# reported here rather than quietly dropped by the join.
bayes_si_checked <- reactive({
  req(bayes_si_global())
  d <- bayes_si_global()
  if (input$sex_bayes_si != "All" && "sex" %in% names(d)) {
    d <- d %>% filter(sex == input$sex_bayes_si)
  }
  if (input$age_bayes_si != "All" && "age_group" %in% names(d)) {
    d <- d %>% filter(age_group == input$age_bayes_si)
  }
  
  full <- bayes_si_global()
  i <- match(d$source_clean, full$id_clean)
  d$onset_read          <- d$onset_min
  d$source_onset_read   <- full$onset_min[i]
  d$serial_days         <- as.numeric(d$onset_min - full$onset_min[i])
  known <- !is.na(d$source_clean) & nzchar(d$source_clean) & d$source_clean %in% full$id_clean
  d$status <- dplyr::case_when(
    is.na(d$onset_min)                             ~ "excluded: onset date could not be read",
    is.na(d$source_clean) | !nzchar(d$source_clean) ~ "excluded: no source of infection given",
    !known                                          ~ "excluded: the source of infection is not one of the case ids in this file",
    is.na(full$onset_min[i])                        ~ "excluded: the source case has no readable onset date",
    TRUE                                            ~ "used"
  )
  d
})

# modeling
bayes_si_fit <- eventReactive(input$run_bayes_si, {
  
  shinyjs::disable("run_bayes_si")
  
  on.exit({
    shinyjs::enable("run_bayes_si")
  })
  
  # cleanup
  gc()
  gc()
  
  req(bayes_si_data())
  df <- bayes_si_data()
  
  validate(
    need(nrow(df) >= 5, "insufficient data: you need at least 5 linked pairs.")
  )
  
  stan_data <- list(
    N = nrow(df),
    tStartExposure = df$tStartExposure,
    tEndExposure = df$tEndExposure,
    tSymptomOnset = df$tSymptomOnset
  )
  
  withProgress(message = 'running bayesian model...', detail = 'initializing...', value = 0, {
    
    mod <- get_compiled_stan_model(input$bayes_model_select_si)
    incProgress(0.1, detail = "sampling chains...")
    
    # run the sampling
    fit <- rstan::sampling(
      object = mod, 
      data = stan_data,
      chains = 4,
      cores = min(4, parallel::detectCores(), na.rm = TRUE),
      seed = STAN_SEED,
      iter = input$iter_count_si, 
      warmup = input$iter_count_si / 2,
      refresh = 0,
      # optimization: only save specific parameters to save ram
      pars = c("alphaInc", "sigmaInc"),
      control = list(adapt_delta = 0.95)
    )
    
    incProgress(0.8, detail = "extracting posteriors...")
    params <- rstan::extract(fit)
    
    # final cleanup after heavy processing
    gc()
    
    list(
      fit_obj = fit,
      iterations = input$iter_count_si,
      model_name = input$bayes_model_select_si,
      shift = if ("si_shift" %in% names(df)) df$si_shift[1] else 0,
      alpha = params$alphaInc,
      sigma = params$sigmaInc,
      alpha_summary = quantile(params$alphaInc, probs = c(0.025, 0.5, 0.975)),
      sigma_summary = quantile(params$sigmaInc, probs = c(0.025, 0.5, 0.975))
    )
  })
})

```

## Row {data-height="180"}

### Sample Size

```{r si-bayes-sample-size}

renderUI({
  req(bayes_si_data())
  
  ss <- nrow(bayes_si_data())
  
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 40px; font-weight: bold; margin-bottom: 5px; text-align: center; font-family: var(--epiparam-font); color: #000000;",
      ss
    ),
    htmltools::div(
      style = "font-size: 14px; text-align: center; color: #666;",
      "Linked Pairs"
    )
  )
})

```

### Model Statistics (Median & 95% CrI)

```{r si-bayes-model-stats}

renderDT({
  req(bayes_si_fit())
  res <- bayes_si_fit()
  
  # dynamic parameter labels
  if (res$model_name == "Lognormal") {
    p_names <- c("MeanLog (Alpha)", "SDLog (Sigma)")
    post_means <- exp(res$alpha + (res$sigma^2)/2)
  } else if (res$model_name == "Gamma") {
    p_names <- c("Shape (Alpha)", "Rate (Sigma)")
    post_means <- res$alpha / res$sigma
  } else if (res$model_name == "Normal") {
    p_names <- c("Mean (Alpha)", "SD (Sigma)")
    post_means <- res$alpha # mean is just alpha for normal
  } else {
    p_names <- c("Shape (Alpha)", "Scale (Sigma)")
    post_means <- res$sigma * gamma(1 + 1/res$alpha)
  }
  
  df_stats <- data.frame(
    Parameter = p_names,
    `Median` = round(c(res$alpha_summary[2], res$sigma_summary[2]), 3),
    `Lower_95_CrI` = round(c(res$alpha_summary[1], res$sigma_summary[1]), 3),
    `Upper_95_CrI` = round(c(res$alpha_summary[3], res$sigma_summary[3]), 3)
  )
  
  # the model was fitted to serial intervals moved to the right by res$shift,
  # so the mean is moved back before it is shown
  si_shift <- if (!is.null(res$shift)) res$shift else 0
  mean_summ <- quantile(post_means - si_shift, probs = c(0.025, 0.5, 0.975))
  
  df_mean <- data.frame(
    Parameter = "Mean Serial Interval",
    `Median` = round(mean_summ[2], 3),
    `Lower_95_CrI` = round(mean_summ[1], 3),
    `Upper_95_CrI` = round(mean_summ[3], 3)
  )
  
  final_df <- rbind(df_stats, df_mean)
  
  # make the shift visible: the two parameters above describe the shifted
  # distribution, the mean has been moved back on to the observed scale
  if (si_shift > 0) {
    final_df <- rbind(final_df, data.frame(
      Parameter = "Shift applied (days)",
      Median = si_shift, Lower_95_CrI = NA, Upper_95_CrI = NA
    ))
  }
  
  fname <- paste0("Serial_Interval_Bayesian_Stats_", Sys.Date())
  
  datatable(
    final_df, 
    rownames = FALSE, 
    extensions = 'Buttons',
    options = list(
      dom = 'Brt',
      buttons = list(
        list(extend = "csv", filename = fname),
        list(extend = "excel", filename = fname),
        list(extend = "pdf", filename = fname)
      ),
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
  ) %>%
    formatStyle(columns = "Parameter", fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Dataset Preview {.data-preview}

```{r si-bayes-dataset-preview}

renderUI({
  d <- bayes_si_checked()
  full <- bayes_si_global()
  known <- !is.na(d$source_clean) & nzchar(d$source_clean) & d$source_clean %in% full$id_clean
  data_check_panel(
    n_read = nrow(d),
    n_used = sum(d$status == "used"),
    problems = problems_from_status(d$status, list(
      "excluded: onset date could not be read" =
        unreadable_values(d$date_of_onsetmin, d$onset_min),
      "excluded: the source of infection is not one of the case ids in this file" =
        unique(d$source_of_infection[!known & !is.na(d$source_clean) & nzchar(d$source_clean)])
    ))
  )
})

renderDataTable({
  # the pairs as the app resolved them, with the status of each row
  dplyr::relocate(bayes_si_checked(), status, id_clean, source_clean,
                  onset_read, source_onset_read, serial_days)
}, options = list(
  pageLength = 10,
  scrollX = TRUE,
  paging = TRUE,
  striped = TRUE,
  bordered = TRUE,
  hover = TRUE
), rownames = FALSE, fillContainer = FALSE)

```

### Posterior Distribution (Median Fit)

```{r si-bayes-posterior-dist}

renderHighchart({
  req(bayes_si_fit())
  res <- bayes_si_fit()
  
  alpha_est <- median(res$alpha)
  sigma_est <- median(res$sigma)
  si_shift  <- if (!is.null(res$shift)) res$shift else 0
  
  # dynamic x-axis range logic
  if (res$model_name == "Normal") {
    # calculate lower bound to capture negative tail (99.9% coverage)
    lower_bound <- qnorm(0.001, mean = alpha_est, sd = sigma_est)
    # cap upper bound at 30 for comparability
    upper_bound <- 30
    x_vals <- seq(lower_bound, upper_bound, length.out = 300)
    x_min_limit <- NULL # let chart auto-scale to show negatives
  } else {
    # strictly positive models are fitted on the shifted scale, so the density
    # is calculated there and drawn back on the observed scale below
    x_vals <- seq(0, 30 + si_shift, length.out = 300)
    x_min_limit <- if (si_shift > 0) NULL else 0
  }
  
  # dynamic density calculation
  if (res$model_name == "Gamma") {
    y_vals <- dgamma(x_vals, shape = alpha_est, rate = sigma_est)
  } else if (res$model_name == "Lognormal") {
    y_vals <- dlnorm(x_vals, meanlog = alpha_est, sdlog = sigma_est)
  } else if (res$model_name == "Normal") {
    y_vals <- dnorm(x_vals, mean = alpha_est, sd = sigma_est)
  } else {
    y_vals <- dweibull(x_vals, shape = alpha_est, scale = sigma_est)
  }
  
  model_label <- tools::toTitleCase(res$model_name)
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  axis_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold", fontFamily = "Jost")
  
  highchart() %>%
    hc_chart(
      style = list(fontFamily = "Jost"), 
      backgroundColor = "#FFFFFF"
    ) %>%
    
    hc_title(
      text = paste0("Posterior Predictive Distribution (", model_label, ")"),
      style = list(fontSize = "18px", fontWeight = "bold", fontFamily = "Jost", color = "#000000")
    ) %>%
    
    hc_subtitle(
      text = paste0("Based on ", res$iterations, " MCMC iterations"),
      style = list(fontSize = "14px", fontFamily = "Jost", color = "#666666")
    ) %>%
    
    hc_exporting(
      enabled = TRUE, 
      filename = paste0("Serial_Interval_Posterior_Distribution_", Sys.Date()),
      buttons = list(contextButton = list(menuItems = menu_items))
    ) %>%
    
    hc_add_series(
      data = data.frame(x = x_vals - si_shift, y = y_vals), 
      type = "area", 
      name = "Posterior Density", 
      color = "rgba(255, 140, 0, 0.5)",
      lineWidth = 2
    ) %>%
    
    hc_xAxis(
      title = list(text = "Serial Interval (Days)", style = axis_style), 
      labels = list(style = axis_style),
      min = x_min_limit,
      minPadding = 0,
      startOnTick = TRUE,
      lineColor = "#000000", 
      lineWidth = 1, 
      tickColor = "#000000"
    ) %>%
    
    hc_yAxis(
      title = list(text = "Density", style = axis_style), 
      labels = list(style = axis_style),
      min = 0, 
      minPadding = 0, 
      startOnTick = TRUE,
      gridLineWidth = 0, 
      lineColor = "#000000", 
      lineWidth = 1, 
      tickColor = "#000000"
    ) %>%
    
    hc_tooltip(valueDecimals = 4, style = list(fontFamily = "Jost"))
})

```

### Trace Plot (Convergence Check)

```{r si-bayes-trace-plot}

renderPlotly({
  req(bayes_si_fit())
  
  p <- rstan::traceplot(bayes_si_fit()$fit_obj, 
                        pars = c("alphaInc", "sigmaInc"),
                        linewidth = 0.25) +
    theme_bw() +
    ggtitle("Trace Plot: Alpha & Sigma") + 
    theme(
      text = element_text(family = "Jost"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10, color = "black"),
      axis.line = element_line(color = "black"),
      strip.background = element_rect(fill = "#f0f0f0"),
      strip.text = element_text(face = "bold")
    )
  
  ggplotly(p) %>%
    config(
      toImageButtonOptions = list(
        format = "png",
        filename = paste0("Serial_Interval_Trace_Plot_", Sys.Date()),
        height = 500,
        width = 700,
        scale = 2
      )
    ) %>%
    layout(font = list(family = "Jost"))
})

```


$R_t$ {data-orientation=rows}
=========================================

Sidebar {.sidebar}
---------------------

```{r rt-sidebar}
hr()
selectInput("rt_source", "Incidence Data",
            choices = c("Upload a file" = "upload",
                        "File loaded on the Incubation Period tab" = "inc",
                        "File loaded on the Serial Interval tab" = "si"),
            selected = "upload")

conditionalPanel(
  condition = "input.rt_source == 'upload'",
  fileInput("file1_rt", "Choose CSV File",
            multiple = FALSE,
            accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
)
hr()

selectInput("rt_si_source", "Serial Interval",
            choices = c("Estimated on the Serial Interval tab" = "app",
                        "Enter the mean and SD" = "manual"),
            selected = "app")

conditionalPanel(
  condition = "input.rt_si_source == 'manual'",
  numericInput("rt_mean_si", "Mean Serial Interval (days)", value = 12.0, min = 1.01, step = 0.1),
  numericInput("rt_std_si", "SD of Serial Interval (days)", value = 5.2, min = 0.1, step = 0.1)
)
hr()

sliderInput("rt_window", "Sliding Window (days)", value = 7, min = 2, max = 21, step = 1)

hr()

valueBox(
  value = downloadButton("downloadMockRt", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

hr()

valueBox(
  value = tags$h6("Note: Rt is estimated with the method of Cori et al. (2013), using the EpiEstim package, over sliding windows of the length set above. Incidence is counted by date of onset. The serial interval is taken from the Serial Interval tab or entered by hand; the manual defaults of 12.0 and 5.2 days suit the mock data.",
                  style = "text-align: justify; font-family: var(--epiparam-font); font-size: 16px; color: white;"),
  color = "#FFFFFF")

hr()
# logo image
renderImage({
  list(
    src = "one_logo.png",
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "display: block; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = FALSE)

```


```{r rt-data-processing}

# the Rt tab estimates from a file of its own, or from the file already loaded on
# the incubation period or serial interval tab, so one outbreak dataset can serve
# all three analyses
rt_file <- reactive({
  src <- input$rt_source
  f <- switch(src,
              inc    = input$file1_inc,
              si     = input$file1_si,
              upload = input$file1_rt)
  req(f)
  d <- read.csv(f$datapath[1], stringsAsFactors = FALSE)
  d <- data.frame(lapply(d, trimws), stringsAsFactors = FALSE)
  clean_names(d)
})

# every row, with the onset date as the app read it and a note of what became of
# the row. the incubation and serial interval files call the column date_of_onset,
# the mock data calls it date_onset; both are accepted.
rt_checked <- reactive({
  d <- rt_file()
  col <- intersect(c("date_of_onset", "date_onset"), names(d))[1]
  validate(need(!is.na(col),
    "incorrect dataset: the Rt tab needs a 'date of onset' (or 'date onset') column."))
  raw   <- as.character(d[[col]])
  blank <- is.na(raw) | !nzchar(trimws(raw)) |
           toupper(trimws(raw)) %in% c("NA", "N/A", "NULL", "-", "ASYMPTOMATIC")
  d$onset_as_given <- raw
  d$onset_read     <- parse_date(raw)
  d$status <- dplyr::case_when(
    blank               ~ "excluded: no onset date given",
    is.na(d$onset_read) ~ "excluded: onset date could not be read",
    TRUE                ~ "used"
  )
  d
})

# daily incidence by date of onset, with the days that had no case filled in as zero
rt_incidence <- reactive({
  d <- rt_checked()
  dates <- d$onset_read[d$status == "used"]
  validate(need(length(dates) > 0, "no usable onset dates in this file."))
  inc <- tibble::tibble(dates = dates) %>%
    dplyr::count(dates, name = "I") %>%
    tidyr::complete(dates = seq(min(dates), max(dates), by = "day"), fill = list(I = 0))
  as.data.frame(inc)
})

# the serial interval behind the estimate: the fit already made on the serial
# interval tab, as the reviewer asked, or a mean and sd typed in by the user
rt_si <- reactive({
  if (identical(input$rt_si_source, "manual")) {
    return(list(label = "entered by hand", mean = input$rt_mean_si, sd = input$rt_std_si))
  }
  req(input$file1_si)
  res <- fit_results_si()
  validate(need(!is.null(res) && length(res) > 0,
    "the Serial Interval tab could not fit a model to its data, so there is no estimate to use."))
  # the model chosen on that tab; when it shows all of them, the best fit by AIC
  by_aic <- length(res) > 1
  best <- if (!by_aic) res[[1]] else
    res[[which.min(vapply(res, function(x) x$aic, numeric(1)))]]
  list(label = paste0(tools::toTitleCase(tolower(best$model)), " fit",
                      if (by_aic) " (lowest AIC)" else "", ", Serial Interval tab"),
       mean  = unname(best$moments["mean"]),
       sd    = sqrt(unname(best$moments["var"])))
})

# Rt by the method of Cori et al. (2013): EpiEstim with a parametric serial interval
rt_result <- reactive({
  inc <- rt_incidence()
  si  <- rt_si()
  w   <- input$rt_window
  validate(
    need(is.numeric(si$mean) && length(si$mean) == 1 && is.finite(si$mean) && si$mean > 1,
         sprintf("the method of Cori et al. needs a mean serial interval above 1 day; the value in use is %s.",
                 format(round(as.numeric(si$mean), 2)))),
    need(is.numeric(si$sd) && length(si$sd) == 1 && is.finite(si$sd) && si$sd > 0,
         "the standard deviation of the serial interval must be above 0."),
    need(nrow(inc) >= w + 1,
         sprintf("a %d-day window needs at least %d days of incidence, and this data covers %d. Shorten the window or use a longer series.",
                 w, w + 1, nrow(inc)))
  )
  # windows of w days sliding one day at a time, starting on day 2 as EpiEstim requires
  t_start <- seq(2, nrow(inc) - w + 1)
  t_end   <- t_start + w - 1
  warn <- character(0)
  res <- withCallingHandlers(
    EpiEstim::estimate_R(
      incid  = inc,
      method = "parametric_si",
      config = EpiEstim::make_config(list(mean_si = si$mean, std_si = si$sd,
                                          t_start = t_start, t_end = t_end))
    ),
    warning = function(x) { warn <<- c(warn, conditionMessage(x)); invokeRestart("muffleWarning") },
    message = function(m) invokeRestart("muffleMessage")
  )
  est <- res$R
  list(
    res = res, si = si, inc = inc, window = w,
    table = data.frame(
      window_start = res$dates[est$t_start],
      window_end   = res$dates[est$t_end],
      mean_rt      = est[["Mean(R)"]],
      median_rt    = est[["Median(R)"]],
      lower_95     = est[["Quantile.0.025(R)"]],
      upper_95     = est[["Quantile.0.975(R)"]]
    ),
    warnings = unique(warn)
  )
})

rt_box <- function(value) {
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: var(--epiparam-font);",
      value
    )
  )
}

```


## Row {data-height="180"}

### Cases

```{r rt-cases}

renderUI({
  inc <- rt_incidence()
  rt_box(format(sum(inc$I), big.mark = ","))
})

```

### Serial Interval

```{r rt-si-used}

renderUI({
  si <- rt_si()
  rt_box(sprintf("%.2f days", si$mean))
})

```

### Latest Rt

```{r rt-latest}

renderUI({
  r <- rt_result()
  last <- utils::tail(r$table, 1)
  rt_box(sprintf("%.2f", last$mean_rt))
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Rt Over Time

```{r rt-plot}

renderHighchart({
  r <- rt_result()
  t <- r$table
  axis_style  <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items  <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  ts <- highcharter::datetime_to_timestamp(t$window_end)
  band <- lapply(seq_len(nrow(t)), function(i) list(x = ts[i], low = t$lower_95[i], high = t$upper_95[i]))
  line <- lapply(seq_len(nrow(t)), function(i) list(x = ts[i], y = t$mean_rt[i]))

  hc <- highchart() %>%
    hc_chart(style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
    hc_title(text = "Effective Reproduction Number (Rt)",
             style = list(fontSize = "16px", fontWeight = "bold")) %>%
    hc_xAxis(type = "datetime", title = list(text = "Date (last day of window)", style = axis_style),
             labels = list(style = label_style), lineColor = "#000000", lineWidth = 1, tickColor = "#000000") %>%
    hc_yAxis(min = 0, title = list(text = "Rt", style = axis_style), labels = list(style = label_style),
             plotLines = list(list(value = 1, color = "#C0392B", width = 1.5, dashStyle = "Dash", zIndex = 5,
                                   label = list(text = "Rt = 1", align = "right",
                                                style = list(color = "#C0392B", fontWeight = "bold"))))) %>%
    hc_add_series(type = "arearange", name = "95% credible interval", data = band,
                  color = "#8B5A00", fillOpacity = 0.22, lineWidth = 0, marker = list(enabled = FALSE)) %>%
    hc_add_series(type = "line", name = "Mean Rt", data = line,
                  color = "#8B5A00", lineWidth = 2.5, marker = list(enabled = FALSE)) %>%
    hc_tooltip(shared = TRUE, valueDecimals = 2, xDateFormat = "%d %b %Y") %>%
    hc_exporting(enabled = TRUE, filename = paste0("Rt_", Sys.Date()),
                 buttons = list(contextButton = list(menuItems = menu_items)))
  hc
})

```

### Incidence

```{r rt-incidence-plot}

renderHighchart({
  inc <- rt_incidence()
  axis_style  <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items  <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  ts  <- highcharter::datetime_to_timestamp(inc$dates)
  pts <- lapply(seq_len(nrow(inc)), function(i) list(x = ts[i], y = inc$I[i]))

  highchart() %>%
    hc_chart(type = "column", style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
    hc_title(text = "Daily Incidence by Date of Onset", style = list(fontSize = "16px", fontWeight = "bold")) %>%
    hc_xAxis(type = "datetime", title = list(text = "Date of onset", style = axis_style),
             labels = list(style = label_style), lineColor = "#000000", lineWidth = 1, tickColor = "#000000") %>%
    hc_yAxis(min = 0, allowDecimals = FALSE, title = list(text = "Cases", style = axis_style),
             labels = list(style = label_style)) %>%
    hc_plotOptions(column = list(pointPadding = 0, groupPadding = 0, borderWidth = 0.2, borderColor = "#FFFFFF")) %>%
    hc_add_series(type = "column", name = "Cases", data = pts, color = "#DAA520") %>%
    hc_legend(enabled = FALSE) %>%
    hc_tooltip(xDateFormat = "%d %b %Y") %>%
    hc_exporting(enabled = TRUE, filename = paste0("Incidence_", Sys.Date()),
                 buttons = list(contextButton = list(menuItems = menu_items)))
})

```

### Serial Interval Distribution

```{r rt-si-plot}

renderHighchart({
  r <- rt_result()
  p <- as.numeric(r$res$si_distr)
  day <- seq_along(p) - 1
  # show the distribution up to the day by which almost all of its mass has fallen
  last_day <- max(which(cumsum(p) <= 0.999)) + 1
  last_day <- min(max(last_day, 2), length(p))
  axis_style  <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items  <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")

  highchart() %>%
    hc_chart(type = "column", style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF") %>%
    hc_title(text = "Serial Interval Used for Rt", style = list(fontSize = "16px", fontWeight = "bold")) %>%
    hc_xAxis(categories = as.character(day[seq_len(last_day)]),
             title = list(text = "Serial interval (days)", style = axis_style), labels = list(style = label_style)) %>%
    hc_yAxis(min = 0, title = list(text = "Probability", style = axis_style), labels = list(style = label_style)) %>%
    hc_plotOptions(column = list(pointPadding = 0.05, groupPadding = 0, borderWidth = 0)) %>%
    hc_add_series(type = "column", name = "Probability", data = round(p[seq_len(last_day)], 4), color = "#228B22") %>%
    hc_legend(enabled = FALSE) %>%
    hc_exporting(enabled = TRUE, filename = paste0("Serial_Interval_Used_", Sys.Date()),
                 buttons = list(contextButton = list(menuItems = menu_items)))
})

```

### Estimates

```{r rt-table}

renderDT({
  t <- rt_result()$table
  out <- data.frame(
    `Window start`  = format(t$window_start, "%Y-%m-%d"),
    `Window end`    = format(t$window_end, "%Y-%m-%d"),
    `Mean Rt`       = round(t$mean_rt, 3),
    `Median Rt`     = round(t$median_rt, 3),
    `Lower 95% CrI` = round(t$lower_95, 3),
    `Upper 95% CrI` = round(t$upper_95, 3),
    check.names = FALSE
  )
  fname <- paste0("Rt_estimates_", Sys.Date())
  datatable(
    out, rownames = FALSE, extensions = "Buttons",
    options = list(
      dom = "Bfrtip",
      buttons = list(list(extend = "csv", filename = fname),
                     list(extend = "excel", filename = fname),
                     list(extend = "pdf", filename = fname)),
      pageLength = 15,
      # EpiEstim gives no estimate for windows that end before the mean serial
      # interval; say so rather than leaving the cells empty
      columnDefs = list(list(targets = 2:5, render = DT::JS(
        "function(data, type) { return (data === null && type === 'display') ? 'not estimable' : data; }"
      )))
    )
  ) %>%
    formatStyle(columns = names(out), `font-size` = "12px")
}, server = FALSE)   # the whole table goes to the browser, so CSV/Excel/PDF export every window, not just the page on screen

```

### Dataset Preview {.data-preview}

```{r rt-data-check}

renderUI({
  d <- rt_checked()
  data_check_panel(
    n_read = nrow(d),
    n_used = sum(d$status == "used"),
    problems = problems_from_status(d$status, list(
      "excluded: onset date could not be read" =
        unique(d$onset_as_given[d$status == "excluded: onset date could not be read"])
    ))
  )
})

```

```{r rt-dataset-preview}

renderDataTable({
  # the onset date as the app read it comes first, with what became of each row
  dplyr::relocate(rt_checked(), status, onset_read)
}, options = list(
  pageLength = 10,
  scrollX = TRUE,
  paging = TRUE,
  striped = TRUE,
  bordered = TRUE,
  hover = TRUE
), rownames = FALSE, fillContainer = FALSE)

```


$k$ and $R_e$ {data-orientation=rows}
=======================================

Sidebar {.sidebar}
-------------------

```{r}

hr()
fileInput("file1_kr", "Choose CSV File",
          multiple = TRUE,
          accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
hr()
# cluster selection for transmission network visualization - initialize with placeholder
selectInput(inputId = "cluster_kr", label = "Select cluster:",
            choices = c("All"), selected = "All")
hr()
valueBox(
  value = tags$h6("Note: Superspreading event is an event in which an infectious disease is spread much more than usual. Superspreader: An unusually contagious individual infected with a disease. The superspreading potential is estimated through the R₀ (Basic Reproduction number) and Kappa (variation in individual infectiousness).", 
                  style = "text-align: justify; font-family: var(--epiparam-font); font-size: 16px; color: white;"),
  color = "#FFFFFF"
)
hr()
valueBox(
  value = downloadButton("downloadMockKappa", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

# logo image
renderImage({
  list(
    src = "one_logo.png",
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "display: block; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = FALSE)

```

```{r k-and-re-data-processing}

# reactive data for k and re - filtered by selected cluster
dataset_kr <- eventReactive(input$file1_kr, {
  req(input$file1_kr)
  dataset <- read.csv(input$file1_kr$datapath)
  dataset <- data.frame(lapply(dataset, trimws))
  dataset <- janitor::clean_names(dataset)
  
  # validation start
  validate(
    need(all(c("cluster_id", "transmission_relationship") %in% names(dataset)),
         "incorrect dataset: please upload k and re data containing 'cluster id' and 'transmission relationship'.")
  )
  # validation end
  
  dataset
})

# reactive dataset filtered by selected cluster
filtered_dataset_kr <- reactive({
  req(dataset_kr(), input$cluster_kr)
  dataset <- dataset_kr()
  
  # filter data based on selected cluster
  if (input$cluster_kr != "All") {
    dataset <- dataset %>% filter(cluster_id == input$cluster_kr)
  }
  
  dataset
})

# update cluster choices when data loads
observeEvent(dataset_kr(), {
  req(dataset_kr())
  clusters <- c("All", unique(dataset_kr()$cluster_id) %>% sort())
  updateSelectInput(session, "cluster_kr", choices = clusters, selected = "All")
})

# helper function for offspring distribution - uses filtered data
get_offspring_data <- function() {
  req(filtered_dataset_kr())
  dt <- filtered_dataset_kr()
  if (nrow(dt) == 0) stop("no data available for selected cluster")
  freqt <- table(dt$source_of_infection_id) %>% data.frame() %>% filter(Var1 != 'unknown')
  zeros <- nrow(dt) - nrow(freqt)
  offt <- table(freqt$Freq) %>% data.frame()
  offtk <- c(rep(0, zeros), rep(as.numeric(levels(offt$Var1)), offt$Freq))
  list(offtk = offtk, fit = fitdistr(offtk, "negative binomial"))
}

```

## Row {data-height="180"}

### Sample Size

```{r}

renderUI({
  # ensure filtered dataset is available
  req(filtered_dataset_kr())
  
  # data Preparation
  ss <- nrow(filtered_dataset_kr())
  
  # render the sample size text
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: var(--epiparam-font);",
      ss
    ),
    htmltools::div(
      style = "font-size: 14px; text-align: center; color: #666;",
      paste("Cluster:", input$cluster_kr)
    )
  )
})

```

### Model Parameters

```{r}

renderDT({
  req(input$file1_kr)
  
  res <- tryCatch({
    get_offspring_data()
  }, error = function(e) {
    NULL
  })
  
  if (is.null(res) || is.null(res$fit)) {
    df <- data.frame(
      parameter = c("Rₑ", "Kappa"),
      value = rep("No data available", 2),
      check.names = FALSE  
    )
  } else {
    fit_df <- as.data.frame(unlist(res$fit))
    fit_df$parameter <- row.names(fit_df)
    rownames(fit_df) <- NULL
    colnames(fit_df) <- c("value", "parameter")
    
    fit_df <- dplyr::select(fit_df, parameter, value) %>%
      mutate(
        parameter = gsub("size", "Kappa", parameter),
        parameter = gsub("mu", "Rₑ", parameter),
        value = round(value, 3)
      ) %>%
      head(4)
    
    df <- fit_df
  }
  
  datatable(
    df,
    rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list(
        list(extend = "csv", filename = paste0("k_ro_model_params_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "excel", filename = paste0("k_ro_model_params_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "pdf", filename = paste0("k_ro_model_params_", input$cluster_kr, "_", Sys.Date()))
      ),
      pageLength = nrow(df)
    ),
    extensions = "Buttons",
    class = "compact cell-border stripe"
  ) %>%
    formatStyle(
      columns = names(df),
      `font-size` = "14px"
    )
})

```

### Confidence Interval

```{r}

renderDT({
  req(input$file1_kr)
  res <- tryCatch({
    get_offspring_data()
  }, error = function(e) {
    NULL
  })
  
  if (is.null(res) || is.null(res$fit)) {
    df <- data.frame(
      parameter = c("Rₑ", "Kappa"),
      `2.5 %` = rep("No data available", 2),
      `97.5 %` = rep("No data available", 2),
      check.names = FALSE  
    )
  } else {
    conf_interval <- confint(res$fit, level = 0.95)
    conf_df <- as.data.frame(conf_interval)
    conf_df$parameter <- rownames(conf_df)
    
    conf_df <- dplyr::select(conf_df, parameter, `2.5 %`, `97.5 %`) %>%
      mutate(
        parameter = gsub("size", "Kappa", parameter),
        parameter = gsub("mu", "Rₑ", parameter),
        `2.5 %` = round(`2.5 %`, 3),
        `97.5 %` = round(`97.5 %`, 3)
      )
    
    df <- conf_df
  }
  
  datatable(
    df,
    rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list(
        list(extend = "csv", filename = paste0("k_ro_conf_int_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "excel", filename = paste0("k_ro_conf_int_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "pdf", filename = paste0("k_ro_conf_int_", input$cluster_kr, "_", Sys.Date()))
      ),
      pageLength = nrow(df)
    ),
    extensions = "Buttons",
    class = "compact cell-border stripe"
  ) %>%
    formatStyle(
      columns = names(df),
      `font-size` = "14px"
    )
})

```


Column {.tabset .tabset-fade }
-------------------------------

### Histogram

```{r}

renderHighchart({
  
  req(input$file1_kr)
  
  res <- tryCatch({
    get_offspring_data()
  }, error = function(e) {
    NULL
  })
  
  if (is.null(res)) {
    # create empty chart with message
    highchart() %>%
      hc_title(
        text = paste("Only one record available or No data available for cluster:", input$cluster_kr), 
        style = list(
          fontSize = "20px", 
          fontWeight = "bold", 
          color = "#000000", 
          fontFamily = "Jost"
        )
      ) %>%
      hc_chart(style = list(fontFamily = "Jost"), backgroundColor = "#FFFFFF")
  } else {
    # calculate negative binomial distribution
    xvalue <- 0:max(res$offtk)
    yvalue <- dnbinom(xvalue, size = res$fit$estimate["size"], mu = res$fit$estimate["mu"])
    
    # create histogram data
    hist_data <- hist(res$offtk, plot = FALSE, breaks = seq(-0.5, max(res$offtk) + 0.5, by = 1))
    
    # create data for columns
    hist_df <- data.frame(
      from = hist_data$breaks[-length(hist_data$breaks)],
      to = hist_data$breaks[-1],
      y = hist_data$density,
      x = hist_data$mids  # For positioning reference
    )
    
    # style definitions
    axis_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
    label_style <- list(color = "#000000", fontSize = "16px", fontWeight = "Normal")
    
    # create negative binomial curve data (nudged by -0.5)
    fit_data <- data.frame(x = xvalue - 0.5, y = yvalue)
    
    highchart() %>%
      # add histogram bars
      hc_add_series(
        data = list_parse(hist_df),
        type = "column",
        name = "Histogram",
        color = "#F2F2F2",
        backgroundColor = "#FFFFFF",
        borderColor = "#000000",
        borderWidth = 1,
        groupPadding = -0.05,
        pointPadding = 0.05,
        pointRange = 1,
        pointPlacement = 0
      ) %>%
      # add negative binomial curve
      hc_add_series(
        data = fit_data,
        type = "line",
        name = "Negative Binomial Fit",
        color = "#000000",
        lineWidth = 1,
        marker = list(enabled = FALSE)
      ) %>%
      hc_xAxis(
        title = list(text = "Offspring", style = axis_style),
        allowDecimals = FALSE,
        gridLineWidth = 0, 
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = -0.5,
        max = max(res$offtk) + 0.5,
        tickInterval = 1
      ) %>%
      hc_yAxis(
        title = list(text = "Density", style = axis_style),
        gridLineWidth = 0, 
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        allowDecimals = TRUE,
        min = 0
      ) %>%
      hc_title(
        text = paste("Cluster:", input$cluster_kr),
        style = list(fontSize = "16px", fontFamily = "Jost", fontWeight = "bold")
      ) %>%
      hc_tooltip(
        shared = TRUE,
        crosshairs = TRUE,
        valueDecimals = 2,
        headerFormat = "<b>Offspring: {point.x}</b><br/>",
        pointFormat = "<span style='color:#000000'>Density</span>: <b>{point.y:.2f}</b><br/>"
      ) %>%
      hc_exporting(
        enabled = TRUE,
        filename = paste0("k_ro_", input$cluster_kr),
        buttons = list(
          contextButton = list(
            menuItems = list('downloadPNG', 'downloadPDF', 'downloadSVG')
          )
        )
      ) %>%
      hc_plotOptions(
        column = list(
          grouping = FALSE,
          pointRange = 1,
          pointPlacement = 0
        )
      ) %>%
      hc_legend(enabled = FALSE) %>%
      hc_chart(style = list(fontFamily = "Jost"))
  }
})

```

### Transmission relationship

```{r transmission-relationship-output}
visNetworkOutput("network_plot")

```

```{r parse-network-data}

output$download_high_quality <- downloadHandler(
  filename = function() { paste("network_cluster_", input$cluster_kr, ".png", sep = "") },
  content = function(file) {
    temp_html <- tempfile(fileext = ".html")
    visSave(network_plot(), temp_html)
    webshot2::webshot(temp_html, file, vwidth = 3600, vheight = 2800, zoom = 6, delay = 3)
  }
)

# add a new download handler for PDF
output$download_high_quality_pdf <- downloadHandler(
  filename = function() { paste("network_cluster_", input$cluster_kr, ".pdf", sep = "") },
  content = function(file) {
    temp_html <- tempfile(fileext = ".html")
    visSave(network_plot(), temp_html)
    webshot2::webshot(temp_html, file, vwidth = 3600, vheight = 2800, delay = 3)
  }
)

filtered_data <- reactive({
  req(dataset_kr(), input$cluster_kr)
  df <- dataset_kr()
  
  # For transmission relationship, don't filter when "All" is selected
  # as we want to show specific cluster networks
  if (input$cluster_kr != "All") {
    df_filtered <- df %>% filter(cluster_id == input$cluster_kr)
  } else {
    # return empty data frame when "All" is selected
    df_filtered <- df[0, ]  # empty data frame with same structure
  }
  df_filtered
})

network_data <- reactive({
  req(filtered_data())
  df_filtered <- filtered_data()
  
  # Check if data is empty (when "All" is selected)
  if (nrow(df_filtered) == 0) {
    return(list(nodes = data.frame(), edges = data.frame()))
  }
  
  edges_level1_2 <- df_filtered %>%
    dplyr::select(cluster_id, infecteegen1) %>%
    unique() %>%
    rename(from = cluster_id, to = infecteegen1)
  
  edges_level2_3 <- df_filtered %>%
    dplyr::select(infecteegen1, infecteegen2) %>%
    unique() %>%
    rename(from = infecteegen1, to = infecteegen2)
  
  edges_level3_4 <- df_filtered %>%
    dplyr::select(infecteegen2, infecteegen3) %>%
    unique() %>%
    rename(from = infecteegen2, to = infecteegen3)
  
  edges_level4_5 <- df_filtered %>%
    dplyr::select(infecteegen3, infecteegen4) %>%
    unique() %>%
    rename(from = infecteegen3, to = infecteegen4)
  
  edge_list <- rbind(edges_level1_2, edges_level2_3, edges_level3_4, edges_level4_5) %>%
    filter(!from == "" & !to == "" & !is.na(from) & !is.na(to))
  
  all_nodes <- unique(c(edge_list$from, edge_list$to))
  
  node_clusters <- df_filtered %>%
    dplyr::select(id = cluster_id, cluster_id) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen1, cluster_id)) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen2, cluster_id)) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen3, cluster_id)) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen4, cluster_id)) %>%
    filter(!is.na(id), id != "") %>%
    distinct(id, .keep_all = TRUE)
  
  nodes <- data.frame(
    id = all_nodes,
    label = all_nodes,
    title = paste("ID:", all_nodes, "<br>Cluster:", input$cluster_kr),
    color = "#4169E1",
    font = list(color = "white", face = "bold", size = 40),
    shadow = TRUE,
    size = 25
  ) %>%
    left_join(node_clusters, by = "id") %>%
    mutate(group = ifelse(is.na(cluster_id), "Unknown", as.character(cluster_id)))
  
  list(nodes = nodes, edges = edge_list)
})

output$network_plot <- renderVisNetwork({
  req(network_data())
  data <- network_data()
  nodes <- data$nodes
  edge_list <- data$edges
  
  # Check if "All" is selected or no data available
  if (input$cluster_kr == "All" || nrow(edge_list) == 0) {
    nodes <- data.frame(
      id = 1,
      label = "Please select an individual cluster.",
      title = "Transmission relationships are shown for individual clusters only.<br>Select a specific cluster from the dropdown or check the 'Transmission Networks' tab to see the full network.",
      color = "#f0f0f0",
      font = list(color = "black", size = 20, face = "normal")
    )
    return(
      visNetwork(nodes, data.frame(from = numeric(0), to = numeric(0))) %>%
        visOptions(manipulation = FALSE) %>%
        visNodes(shape = "box", size = 80) %>%
        visLayout(randomSeed = 123) %>%
        visPhysics(stabilization = FALSE) %>%
        visInteraction(dragNodes = FALSE, dragView = FALSE, zoomView = FALSE)
    )
  }
  
  visNetwork(nodes, edge_list) %>%
    visEdges(
      arrows = list(to = list(enabled = TRUE, scaleFactor = 1.2)),
      color = list(color = "#6E6E6E", highlight = "#FF0000"),
      smooth = list(enabled = TRUE, type = "dynamic")
    ) %>%
    visNodes(
      shape = "circle",
      size = 90,
      borderWidth = 1,
      borderWidthSelected = 4,
      font = list(size = 24, color = "white", face = "bold")
    ) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = list(from = 5, to = 5), hover = TRUE, algorithm = "hierarchical"),
      nodesIdSelection = list(enabled = TRUE, values = nodes$id, selected = input$cluster_kr, style = 'width: 200px; height: 26px;')
    ) %>%
    visExport(
      type = "png",
      name = paste("transmission_network_cluster", input$cluster_kr),
      label = "Download as PNG",
      background = "#fff",
      style = "position: absolute; top: -400px; right: 10px; z-index: 10;",
      loadDependencies = TRUE
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE,
      tooltipDelay = 0,
      dragView = TRUE,
      zoomView = TRUE
    ) %>%
    visLayout(
      randomSeed = 123,
      improvedLayout = TRUE
    ) %>%
    visPhysics(
      stabilization = list(iterations = 100),
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -50,
        centralGravity = 0.01,
        springLength = 100,
        springConstant = 0.08
      )
    ) %>%
    visLegend(enabled = FALSE) %>%
    visEvents(
      selectNode = "function(nodes) { Shiny.setInputValue('selected_node', nodes); }"
    )
})

observeEvent(input$selected_node, {
  req(input$selected_node, dataset_kr())
  selected_id <- input$selected_node$nodes[[1]]
  df <- dataset_kr()
  selected_cluster <- df %>%
    filter(cluster_id == selected_id |
             infecteegen1 == selected_id |
             infecteegen2 == selected_id |
             infecteegen3 == selected_id |
             infecteegen4 == selected_id) %>%
    pull(cluster_id) %>%
    first()
  if(!is.na(selected_cluster) && selected_cluster != input$cluster_kr) {
    updateSelectInput(session, "cluster_kr", selected = selected_cluster)
  }
})

```

### Transmission Networks

```{r transmission-networks-output}
visNetworkOutput("full_network_plot")
```


```{r parse-transmission-chain}

# function to parse transmission chains
parse_transmission_chain <- function(relationship) {
  # skip if relationship is empty or unknown
  if (is.na(relationship) || relationship == "" || relationship == "unknown") {
    return(data.frame(from = character(), to = character()))
  }
  # split the chain by the arrow symbol (noticed this â†’ in the csv)
  chain <- strsplit(relationship, "→|â†’")[[1]]
  chain <- trimws(chain)  # added this cos some rows have whitespace
  
  # extract immediate relationships (each consecutive pair)
  if (length(chain) >= 2) {
    edges <- data.frame(
      from = chain[1:(length(chain) - 1)],
      to = chain[2:length(chain)]
    )
    return(edges[edges$from != "" & edges$to != "" & edges$from != edges$to, ])
  }
  
  return(data.frame(from = character(), to = character()))
}

output$download_full_network <- downloadHandler(
  filename = function() { paste("full_transmission_network_", Sys.Date(), ".png", sep = "") },
  content = function(file) {
    temp_html <- tempfile(fileext = ".html")
    visSave(full_network_plot(), temp_html)
    webshot2::webshot(temp_html, file, vwidth = 3600, vheight = 2800, zoom = 6, delay = 3)
  }
)

full_network_data <- reactive({
  req(dataset_kr())
  df <- dataset_kr()
  
  edges_list <- lapply(df$transmission_relationship, parse_transmission_chain)
  edges <- do.call(rbind, edges_list)
  
  edges <- edges[complete.cases(edges), ]
  edges <- unique(edges)
  
  # get all unique nodes and sort them numerically
  all_nodes <- unique(c(edges$from, edges$to))
  
  # extract numeric values for sorting
  node_numbers <- as.numeric(gsub("#", "", all_nodes))
  sorted_nodes <- all_nodes[order(node_numbers)]
  
  # create nodes data frame
  nodes <- data.frame(
    id = sorted_nodes,
    label = sorted_nodes,
    title = paste("ID:", sorted_nodes),
    color = "#4169E1",
    font = list(color = "white", face = "bold", size = 50),
    shadow = TRUE,
    size = 30
  )
  
  list(nodes = nodes, edges = edges, sorted_ids = sorted_nodes)
})

output$full_network_plot <- renderVisNetwork({
  req(full_network_data())
  data <- full_network_data()
  nodes <- data$nodes
  edges <- data$edges
  sorted_ids <- data$sorted_ids
  
  if(nrow(edges) == 0) {
    nodes <- data.frame(
      id = 1,
      label = "No transmission data",
      title = "No transmission relationships found in the dataset",
      color = "#f0f0f0",
      font = list(color = "black", size = 24)
    )
    return(
      visNetwork(nodes, data.frame(from = numeric(0), to = numeric(0))) %>%
        visOptions(manipulation = FALSE) %>%
        visNodes(shape = "box", size = 50) %>%
        visLayout(randomSeed = 123)
    )
  }
  
  visNetwork(nodes, edges) %>%
    visEdges(
      arrows = list(to = list(enabled = TRUE, scaleFactor = 1.2)),
      color = list(color = "#6E6E6E", highlight = "#FF0000"),
      smooth = list(enabled = TRUE, type = "dynamic")
    ) %>%
    visNodes(
      shape = "circle",
      size = 30,
      borderWidth = 1,
      borderWidthSelected = 4,
      font = list(size = 20, color = "white", face = "bold")
    ) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = list(
        enabled = TRUE, 
        values = sorted_ids,
        style = 'width: 200px; height: 26px;'
      )
    ) %>%
    visExport(
      type = "png",
      name = paste("full_transmission_network_", Sys.Date()),
      label = "Download as PNG",
      background = "#fff",
      style = "position: absolute; top: -400px; right: 10px;",
      loadDependencies = TRUE
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE,
      tooltipDelay = 0,
      dragView = TRUE,
      zoomView = TRUE
    ) %>%
    visLayout(
      randomSeed = 123,
      improvedLayout = TRUE
    ) %>%
    visPhysics(
      stabilization = list(iterations = 100),
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -50,
        centralGravity = 0.01,
        springLength = 100,
        springConstant = 0.08
      )
    ) %>%
    visLegend(enabled = FALSE)
})

```

### Dataset Preview {.data-preview}

```{r kappa-re-data-check}
renderUI({
  d   <- filtered_dataset_kr()
  rel <- as.character(d$transmission_relationship)
  n_links <- vapply(rel, function(r) nrow(parse_transmission_chain(r)),
                    integer(1), USE.NAMES = FALSE)
  blank <- is.na(rel) | !nzchar(trimws(rel)) | tolower(trimws(rel)) == "unknown"
  # a row with no arrow is an index case, which is normal. a row that looks like
  # it was meant to be a chain but uses some other separator is not, and used to
  # be dropped without a word.
  other_sep <- !blank & n_links == 0 &
    grepl("->|=>|>|,|;|/|\\bto\\b", rel, ignore.case = TRUE)
  
  problems <- list()
  if (any(other_sep)) {
    problems <- c(problems, list(sprintf(
      paste("%d row%s give a transmission relationship the app could not read: chains must be",
            "written with the arrow \u2192, as in \u201c#1 \u2192 #4\u201d (for example: %s)."),
      sum(other_sep), if (sum(other_sep) == 1) "" else "s",
      paste0("\u201c", utils::head(unique(rel[other_sep]), 3), "\u201d", collapse = ", "))))
  }
  if (any(blank)) {
    problems <- c(problems, list(sprintf(
      "%d row%s have no transmission relationship, and contribute no link.",
      sum(blank), if (sum(blank) == 1) "" else "s")))
  }
  data_check_panel(
    n_read = nrow(d),
    n_used = sum(n_links > 0),
    problems = problems,
    used_label = "with an onward transmission link",
    out_label  = "index cases with none"
  )
})
```

```{r kappa-re-dataset-preview}
renderDataTable({
  d <- filtered_dataset_kr()
  # how many onward links the app actually read out of each row
  d$links_read <- vapply(as.character(d$transmission_relationship),
                         function(r) nrow(parse_transmission_chain(r)),
                         integer(1), USE.NAMES = FALSE)
  dplyr::relocate(d, links_read, transmission_relationship)
}, options = list(pageLength = 10, scrollX = TRUE), rownames = FALSE, fillContainer = FALSE)

```
